---
title: "Fire_06_Calibration_ElasticNet"
author: "Femke Keij S2647168"
date: "2023-02-28"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preliminaries
Clear working directory & set random seed:
```{r}
rm(list = ls(all = TRUE))

set.seed(42)
```

Packages used:
```{r}
# for importing / working with tidy data
library(tidyverse)

# for ggplot
library(directlabels)
library(patchwork)
library(ggbeeswarm)

# for fitting the elastic net
library(caret)
library(glmnet)
```

Source files for error computation and plotting functions:
```{r}
source('scripts/Fire model/Fire_ErrorFunctions.R')
```

Read in the data:
```{r}
# complete output data
fire_output <- read_csv('data/raw/fire_output.csv')

# training data
fire_train <- read_csv(
  'data/processed/fire_train.csv')
fire_train$directions <- as.factor(fire_train$directions)

# test data
fire_test <- read_csv(
  'data/processed/fire_test.csv')
fire_test$directions <- as.factor(fire_test$directions)
```

Random samples:
5, 10, 20, 30, 40, 50, 60, 70, and 80 $%$ of the total parameter space.
```{r}
N <- nrow(fire_output)
# range of (training) sample sizes to work with
n <- c(0.05 * N, 0.1 * N, 0.2 * N, 0.3 * N, 0.4 * N, 0.5 * N,
       0.6 * N, 0.7 * N, nrow(fire_train))
```

# Fitting
Here I'm fitting an elastic net regression to predict density and direction.
```{r}
ElasticNetFitting <- function(ticks = FALSE){
  # table to store all output
  length_tib <- sum(nrow(fire_test) * length(n))
  elastic_results <- tibble(direction_true = numeric(length_tib),
                              direction_pred = numeric(length_tib),
                              density_true = numeric(length_tib),
                              density_pred = numeric(length_tib),
                              burn_true = numeric(length_tib),
                              burn_pred = numeric(length_tib),
                              n = numeric(length_tib),
                              ticks_included = numeric(length_tib))
  # vector to store mean density in each training sample
  # (to calculate point prediction performance later on)
  density_means <- numeric(length(n))
  # list to store fitted Elastic Nets
  elastic_fits <- list()

  count <- 1
  
  # for each sample size
  for(i in 1:length(n)){
    # sample training set
    ind_sample <- sample(1:nrow(fire_train), size = n[i])
    fire_train_sub <- fire_train[ind_sample, ]
    # calculate and save mean density in training set
    density_means[i] <- mean(fire_train_sub$density)
    
    if(ticks == TRUE){
      # Elastic Net to predict density when ticks are included
      fit_elastic_density <- train(density ~ burn_percentage +
                                     directions + ticks,
                                   data = fire_train_sub,
                                   method = 'glmnet',
                                   trControl = 
                                     trainControl(method = 'cv', number = 10),
                                   tuneLength = 10)
      # Elastic Net to predict directions when ticks are included
      fit_elastic_directions <- train(directions ~ burn_percentage +
                                        density + ticks,
                                      data = fire_train_sub,
                                      method = 'glmnet',
                                      trControl =
                                        trainControl(method = 'cv',
                                                     number = 10),
                                      tuneLength = 10)
      
    } else{
      # Elastic Net to predict density when ticks are excluded
      fit_elastic_density <- train(density ~ burn_percentage +
                                     directions,
                                   data = fire_train_sub,
                                   method = 'glmnet',
                                   trControl = 
                                     trainControl(method = 'cv', number = 10),
                                   tuneLength = 10)
      # Elastic Net to predict directions when ticks are excluded
      fit_elastic_directions <- train(directions ~ burn_percentage +
                                        density,
                                      data = fire_train_sub,
                                      method = 'glmnet',
                                      trControl =
                                        trainControl(method = 'cv',
                                                     number = 10),
                                      tuneLength = 10)
    }
    # predict density in test data using fitted Elastic Net
    predict_elastic_density <- predict(fit_elastic_density,
                                       newdata = fire_test)
    # predict direction in test data using fitted regression
    predict_elastic_directions <- predict(fit_elastic_directions,
                                          newdata = fire_test)
    
    # for each instance of the test data
    for(j in 1:nrow(fire_test)){
      # find an instance in the training data that matches the directions
      # and density
      ind <- which(fire_train$density == round(predict_elastic_density[j]) &
                     fire_train$directions ==
                     predict_elastic_directions[j])
      # if there are multiple, pick one
      if(length(ind) > 1){
        ind <- sample(ind, size = 1)
        # if there are none, increase density by 1 and see if there
        # is a matching instance in the training data
        # if not, continue until a matching instance is found
      } else if (length(ind) == 0){
        add <- 1
        while(length(ind) != 1){
          if(round(predict_elastic_density[j]) > 50){
            add <- add * -1
          }
          ind <- which(fire_train$density ==
                         (round(predict_elastic_density[j]) + add) &
                     fire_train$directions ==
                     predict_elastic_directions[j])
          if(length(ind) > 1){
            ind <- sample(ind, size = 1)
          }
          add <- abs(add) + 1
        }
      }
      # store all results in table
      elastic_results$direction_true[count] <- fire_test$directions[j]
      elastic_results$direction_pred[count] <-
        predict_elastic_directions[j]
      elastic_results$density_true[count] <- fire_test$density[j]
      elastic_results$density_pred[count] <- predict_elastic_density[j]
      elastic_results$burn_true[count] <- fire_test$burn_percentage[j]
      elastic_results$burn_pred[count] <- fire_train$burn_percentage[ind]
      elastic_results$n[count] <- n[i]
      elastic_results$ticks_included[count] <- if(ticks){'yes'}else{'no'}
      
      count <- count + 1
    }
    # store regression fit
    elastic_fits <- append(elastic_fits, fit_elastic_density)
    elastic_fits <- append(elastic_fits, fit_elastic_directions)
    
    print(paste('finished fitting sample size ', n[i]))
  }
  
  # correct format of direction_true variable
  elastic_results %>%
    mutate(direction_true = replace(direction_true, direction_true == 1, 4),
         direction_true = replace(direction_true, direction_true == 2, 8),
         direction_pred = replace(direction_pred, direction_pred == 1, 4),
         direction_pred = replace(direction_pred, direction_pred == 2, 8)) ->
    elastic_results
  
  # return list with results, density means, and fitted regression lines
  return(list(elastic_results, density_means, elastic_fits))
}
```

```{r}
# fit regressions with ticks
elastic_results_ticks <- ElasticNetFitting(ticks = TRUE)
density_means_ticks <- elastic_results_ticks[[2]]
elastic_fits_ticks <- elastic_results_ticks[[3]]
elastic_results_ticks <- elastic_results_ticks[[1]]

# fit regressions without ticks
elastic_results_noticks <- ElasticNetFitting(ticks = FALSE)
density_means_noticks <- elastic_results_noticks[[2]]
elastic_fits_noticks <- elastic_results_noticks[[3]]
elastic_results_noticks <- elastic_results_noticks[[1]]
```

```{r}
# run error calculations
elastic_errors_ticks <- ComputeErrors(elastic_results_ticks,
                                         ticks = TRUE,
                                         density_means_ticks,
                                         pred_int_included = FALSE)
elastic_errors_noticks <- ComputeErrors(elastic_results_noticks,
                                           ticks = FALSE,
                                           density_means_noticks,
                                           pred_int_included = FALSE)
```

Merge results
```{r}
elastic_results <- rbind(elastic_results_noticks,
                            elastic_results_ticks)
write_csv(elastic_results, 'data/processed/fire_elasticnet_results.csv')

elastic_errors <- rbind(elastic_errors_noticks, elastic_errors_ticks)
write_csv(elastic_errors, 'data/processed/fire_elasticnet_errors.csv')
``` 

# Plots
```{r}
# elastic_results <- read_csv('data/processed/fire_elasticnet_results.csv')
# elastic_errors <- read_csv('data/processed/fire_elasticnet_errors.csv')
```

## Errors
Fix results and errors data frames:
```{r}
elastic_results <- elastic_results %>%
  mutate(n = factor(n),
         direction_true = factor(direction_true),
         direction_pred = factor(direction_pred),
         direction_correct = (direction_pred == direction_true))

elastic_errors <- elastic_errors %>%
  mutate(n = factor(n))
```

Predicted vs. true density (largest sample size only):
```{r}
PlotPredTrueDensity(elastic_results, n = 1584)

ggsave('figures/fire_elasticnet_predtrue_density.pdf')
```
Predicted vs. true burn percentage (largest sample size only):
```{r}
PlotPredTrueBurn(elastic_results, n = 1584)

ggsave('figures/fire_elasticnet_predtrue_burn.pdf')
```
% correctly predicted parameters (density & directions):
```{r}
PlotPercCorrectParams(elastic_errors, n_plot = c(99, 792, 1584))

ggsave('figures/fire_elasticnet_perc_correct.pdf')
```
RMSE, NRMSE and point prediction performance for density parameter:
```{r}
PlotRMSEParams(elastic_errors, n_plot = c(99, 792, 1584))

ggsave('figures/fire_elasticnet_RMSE_density.pdf')
```
RMSE for burn percentage:
```{r}
PlotRMSEOut(elastic_errors, n_plot = c(99, 792, 1584))

ggsave('figures/fire_elasticnet_RMSE_burn.pdf')
```

## Model fit
For density
```{r}
fit_elastic_density_noticks <- train(density ~ burn_percentage + directions,
                                     data = fire_train,
                                     method = 'glmnet',
                                     trControl = trainControl(method = 'cv',
                                                              number = 10),
                                     tuneLength = 10)
# plot(fit_elastic_density_noticks)

predict_elastic_density_noticks <- data.frame(
  burn_percentage = fire_train$burn_percentage,
  predicted_density = predict(fit_elastic_density_noticks),
  directions = fire_train$directions)

p1 <- fire_train %>%
  ggplot(mapping = aes(x = burn_percentage, y = density,
                       colour = directions)) +
  geom_point() +
  geom_line(data = predict_elastic_density_noticks,
            mapping = aes(x = burn_percentage, y = predicted_density,
                          group = directions, colour = directions)) +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
        panel.border = element_rect(colour = 'lightgrey', fill = NA)) +
  labs(x = '% burned trees at last tick',
       y = 'tree density (%)',
       colour = 'number of \n directions') +
  coord_flip() +
  ggtitle('no ticks')

fit_elastic_density_ticks <- train(density ~ burn_percentage +
                                     directions + ticks,
                                   data = fire_train,
                                   method = 'glmnet',
                                   trControl = 
                                     trainControl(method = 'cv', number = 10),
                                   tuneLength = 10)
# plot(fit_elastic_density_ticks)

predict_elastic_density_ticks <- data.frame(
  burn_percentage = fire_train$burn_percentage,
  predicted_density = predict(fit_elastic_density_ticks),
  directions = fire_train$directions)

p2 <- fire_train %>%
  ggplot(mapping = aes(x = burn_percentage, y = density,
                       colour = directions)) +
  geom_point() +
  geom_line(data = predict_elastic_density_ticks,
            mapping = aes(x = burn_percentage, y = predicted_density,
                          group = directions, colour = directions)) +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
        panel.border = element_rect(colour = 'lightgrey', fill = NA)) +
  labs(x = '% burned trees at last tick',
       y = 'tree density (%)',
       colour = 'number of \n directions') +
  coord_flip() +
  ggtitle('ticks')

patch <- p1 + p2 +
    plot_layout(guides = 'collect') +
    plot_annotation(tag_levels = 'A') & 
    theme(plot.tag = element_text(size = 8)) &
    ylab(NULL)
  
wrap_elements(panel = patch) +
  labs(tag = 'tree density (%)') +
  theme(plot.tag = element_text(size = rel(1)),
        plot.tag.position = 'bottom')

ggsave('figures/fire_elasticnet_fit_density.pdf')
```

For directions
```{r}
fit_elastic_directions_noticks <- train(directions ~ burn_percentage + density,
                                     data = fire_train,
                                     method = 'glmnet',
                                     trControl = trainControl(method = 'cv',
                                                              number = 10),
                                     tuneLength = 10)
# plot(fit_elastic_directions_noticks)

predict_elastic_directions_noticks <- data.frame(
  burn_percentage = fire_train$burn_percentage,
  predicted_directions = predict(fit_elastic_directions_noticks),
  true_directions = fire_train$directions,
  density = fire_train$density)

predict_elastic_directions_noticks <- predict_elastic_directions_noticks %>%
  mutate(directions_correct = (predicted_directions == true_directions))

p1 <- predict_elastic_directions_noticks %>%
  ggplot(mapping = aes(x = density, y = burn_percentage,
                       colour = directions_correct)) +
  geom_point() +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
        panel.border = element_rect(colour = 'lightgrey', fill = NA)) +
  labs(x = 'tree density (%)',
       y = '% burned trees at last tick',
       colour = 'number of \n directions \n predicted \n correctly') +
  ggtitle('no ticks')

fit_elastic_directions_ticks <- train(directions ~ burn_percentage + density +
                                        ticks,
                                     data = fire_train,
                                     method = 'glmnet',
                                     trControl = trainControl(method = 'cv',
                                                              number = 10),
                                     tuneLength = 10)
# plot(fit_elastic_directions_ticks)

predict_elastic_directions_ticks <- data.frame(
  burn_percentage = fire_train$burn_percentage,
  predicted_directions = predict(fit_elastic_directions_ticks),
  true_directions = fire_train$directions,
  density = fire_train$density)

predict_elastic_directions_ticks <- predict_elastic_directions_ticks %>%
  mutate(directions_correct = (predicted_directions == true_directions))

p2 <- predict_elastic_directions_ticks %>%
  ggplot(mapping = aes(x = density, y = burn_percentage,
                       colour = directions_correct)) +
  geom_point() +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
        panel.border = element_rect(colour = 'lightgrey', fill = NA)) +
  labs(x = 'tree density (%)',
       y = '% burned trees at last tick',
       colour = 'number of \n directions \n predicted \n correctly') +
  ggtitle('ticks')

patch <- p1 + p2 +
    plot_layout(guides = 'collect') +
    plot_annotation(tag_levels = 'A') & 
    theme(plot.tag = element_text(size = 8)) &
    xlab(NULL)
  
wrap_elements(panel = patch) +
  labs(tag = 'tree density (%)') +
  theme(plot.tag = element_text(size = rel(1)),
        plot.tag.position = 'bottom')

ggsave('figures/fire_elasticnet_fit_directions.pdf')
``` 
