---
title: "Fire_03_Calibration_SimpleRegression"
author: "Femke Keij S2647168"
date: "2023-02-14"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preliminaries
Clear working directory & set random seed:
```{r}
rm(list = ls(all = TRUE))

set.seed(42)
```

Packages used:
```{r}
# for importing / working with tidy data
library(tidyverse)

# for ggplot
library(directlabels)
library(patchwork)
library(ggbeeswarm)
```

Source files for error computation and plotting functions:
```{r}
source('scripts/Fire model/Fire_ErrorPlots.R')
source('scripts/Fire model/Fire_ErrorCalc.R')
```

Read in the data:
```{r}
# complete output data
fire_output <- read_csv('data/raw/fire_output.csv')

# training data
fire_train <- read_csv(
  'data/processed/fire_train.csv')
fire_train$directions <- as.factor(fire_train$directions)

# test data
fire_test <- read_csv(
  'data/processed/fire_test.csv')
fire_test$directions <- as.factor(fire_test$directions)
```

Random samples:
5, 10, 20, 30, 40, 50, 60, 70, and 80 $%$ of the total parameter space.
```{r}
N <- nrow(fire_output)
# range of (training) sample sizes to work with
n <- c(0.05 * N, 0.1 * N, 0.2 * N, 0.3 * N, 0.4 * N, 0.5 * N,
       0.6 * N, 0.7 * N, nrow(fire_train))
```

# Fitting
Fitting regressions to predict density and directions. For density, a simple linear regression with predictors percentage of burned trees, directions of fire spreading, and total number of ticks is used. Since the directions are categorical, a logistic regression with predictors percentage of burned trees, tree density, and total number of ticks is used. These regressions are then used to predict the density and directions based on the burn percentages in the test data.
```{r}
LinearRegressionFitting <- function(ticks = FALSE){
  # table to store all output
  length_tib <- sum(nrow(fire_test) * length(n))
  regression_results <- tibble(direction_true = numeric(length_tib),
                              direction_pred = numeric(length_tib),
                              density_true = numeric(length_tib),
                              density_pred = numeric(length_tib),
                              density_pred_lwr = numeric(length_tib),
                              density_pred_upr = numeric(length_tib),
                              burn_true = numeric(length_tib),
                              burn_pred = numeric(length_tib),
                              n = numeric(length_tib),
                              ticks_included = numeric(length_tib))
  # vector to store mean density in each training sample
  # (to calculate point prediction performance later on)
  density_means <- numeric(length(n))
  # list to store fitted regression lines
  regression_fits <- list()

  count <- 1
  
  # for each sample size
  for(i in 1:length(n)){
    # sample training set
    ind_sample <- sample(1:nrow(fire_train), size = n[i])
    fire_train_sub <- fire_train[ind_sample, ]
    # calculate and save mean density in training set
    density_means[i] <- mean(fire_train_sub$density)
    
    if(ticks == TRUE){
      # linear regression to predict density when ticks are included
      fit_lm_density <- lm(data = fire_train_sub,
                               formula = 'density ~ burn_percentage + directions + ticks')
      # logistic regression to predict directions when ticks are included
      fit_glm_directions <- glm(data = fire_train_sub,
                                    formula = 'directions ~ burn_percentage + density + ticks',
                                    family = binomial(link = 'logit'))
      
    } else{
      # linear regression to predict density when ticks are excluded
      fit_lm_density <- lm(data = fire_train_sub,
                               formula = 'density ~ burn_percentage + directions')
      # logistic regression to predict directions when ticks are excluded
      fit_glm_directions <- glm(data = fire_train_sub,
                                      formula = 'directions ~ burn_percentage + density',
                                      family = binomial(link = 'logit'))
    }
    # predict density in test data using fitted regression
    # include prediction interval to compute coverage later on
    predict_lm_density <- predict.lm(fit_lm_density,
                                             newdata = fire_test,
                                     interval = 'prediction')
    # predict direction in test data using fitted regression
    predict_glm_directions <- predict.glm(fit_glm_directions,
                                                  newdata = fire_test,
                                                  type = 'response')
    # correct the predicted directions from '1' and '2' to '4' and '8'
    # to match test data
    for(k in 1:nrow(fire_test)){
      if(predict_glm_directions[k] < 0.5){
        predict_glm_directions[k] <- 4
      } else{
        predict_glm_directions[k] <- 8
      }
    }
    
    # for each instance of the test data
    for(j in 1:nrow(fire_test)){
      # find an instance in the training data that matches the directions
      # and density
      ind <- which(fire_train$density == round(predict_lm_density[j,1]) &
                     fire_train$directions ==
                     predict_glm_directions[j])
      if(length(ind) > 1){
        ind <- sample(ind, size = 1)
      }
      # find predicted burn percentage based on training data
      burn_pred <- fire_train$burn_percentage[ind]
      
      # store all results in table
      regression_results$direction_true[count] <- fire_test$directions[j]
      regression_results$direction_pred[count] <-
        predict_glm_directions[j]
      regression_results$density_true[count] <- fire_test$density[j]
      regression_results$density_pred[count] <- predict_lm_density[j,1]
      regression_results$density_pred_lwr[count] <-
        predict_lm_density[j,2]
      regression_results$density_pred_upr[count] <-
        predict_lm_density[j,3]
      regression_results$burn_true[count] <- fire_test$burn_percentage[j]
      regression_results$burn_pred[count] <- burn_pred
      regression_results$n[count] <- n[i]
      regression_results$ticks_included[count] <- if(ticks){'yes'}else{'no'}
      
      count <- count + 1
    }
    # store regression fit
    regression_fits <- append(regression_fits, fit_lm_density)
  }
  
  # correct format of direction_true variable
  regression_results %>%
    mutate(direction_true = replace(direction_true, direction_true == 1, 4),
         direction_true = replace(direction_true, direction_true == 2, 8)) ->
    regression_results
  
  # return table with results, density means, and fitted regression lines
  return(list(regression_results, density_means, regression_fits))
}
```

```{r}
# fit regressions with ticks
regression_results_ticks <- LinearRegressionFitting(ticks = TRUE)
density_means_ticks <- regression_results_ticks[[2]]
regression_fits_ticks <- regression_results_ticks[[3]]
regression_results_ticks <- regression_results_ticks[[1]]

# fit regressions without ticks
regression_results_noticks <- LinearRegressionFitting(ticks = FALSE)
density_means_noticks <- regression_results_noticks[[2]]
regression_fits_noticks <- regression_results_noticks[[3]]
regression_results_noticks <- regression_results_noticks[[1]]
```

```{r}
# run error calcualtions
regression_errors_ticks <- ComputeErrors(regression_results_ticks,
                                         ticks = TRUE,
                                         density_means_ticks,
                                         pred_int_included = TRUE)
regression_errors_noticks <- ComputeErrors(regression_results_noticks,
                                           ticks = FALSE,
                                           density_means_noticks,
                                           pred_int_included = TRUE)
```

Merge results
```{r}
regression_results <- rbind(regression_results_noticks,
                            regression_results_ticks)

regression_errors <- rbind(regression_errors_noticks, regression_errors_ticks)
``` 

# Plots
Fix results and errors data frames:
```{r}
regression_results$n <- as.factor(regression_results$n)
regression_errors$n <- as.factor(regression_errors$n)
regression_results$direction_true <-
  as.factor(regression_results$direction_true)
regression_results$direction_pred <-
  as.factor(regression_results$direction_pred)
regression_results %>%
  mutate(direction_correct = (direction_pred == direction_true)) -> regression_results
```

Predicted vs. true density (largest sample size only):
```{r}
PlotPredTrueDensity(regression_results, n = 1584)

#ggsave('figures/fire_LHS_pred_vs_true.pdf')
```
Predicted vs. true burn percentage (largest sample size only):
```{r}
PlotPredTrueBurn(regression_results, n = 1584)
```
% correctly predicted parameters (density & directions):
```{r}
PlotPercCorrectParams(regression_errors, n_plot = c(99, 792, 1584))

#ggsave('figures/fire_LHS_perc_correct.pdf')
```
RMSE, NRMSE and point prediction performance for density parameter:
```{r}
PlotRMSEParams(regression_errors, n_plot = c(99, 792, 1584))

#ggsave('figures/fire_LHS_RMSE.pdf')
```
RMSE for burn percentage:
```{r}
PlotRMSEOut(regression_errors, n_plot = c(99, 792, 1584))
```
Coverage:
```{r}
PlotCoverageDensity(regression_errors, n_plot = n)

ggsave('figures/fire_regression_coverage.pdf')
```






Fitted regression lines (smallest and largest sample size):
Linear regression for density
```{r}
RegressionLinePlot <- function(sample = 1, ticks = TRUE){
  if(ticks){
    formula <- 'density ~ burn_percentage + directions + ticks'
    subtitle <- 'with ticks'
  } else{
    formula <- 'density ~ burn_percentage + directions'
    subtitle <- 'without ticks'
  }
  
  subtitle <- paste(subtitle, ' sample size ', n[sample])
  
  ind_sample <- sample(1:nrow(fire_train), size = n[sample])
  fire_train_sub <- fire_train[ind_sample, ]
  
  fire_train_sub %>%
    ggplot(mapping = aes(x = burn_percentage, y = density,
                         colour = directions)) +
    geom_point() +
    geom_line(data = fortify(lm(data = fire_train_sub,
                                formula = formula)),
              aes(x = burn_percentage, y = .fitted)) +
    labs(x = '% burned trees at last tick',
         y = 'tree density (%)',
         subtitle = subtitle) +
    theme_minimal() +
    coord_flip() -> returnplot

  return(returnplot)
}

p1 <- RegressionLinePlot(sample = 1, ticks = TRUE)
p2 <- RegressionLinePlot(sample = 1, ticks = FALSE)
p3 <- RegressionLinePlot(sample = 9, ticks = TRUE)
p4 <- RegressionLinePlot(sample = 9, ticks = FALSE)

p1 + p2 + p3 + p4 +
  plot_layout(guides = 'collect') +
  plot_annotation(tag_levels = 'A') & 
  theme(plot.tag = element_text(size = 8))

ggsave('figures/fire_regression_fit.pdf')
```

Logistic regression for directions
```{r}
library(broom)

fit_glm <- glm(data = fire_train,
               formula = 'directions ~ burn_percentage + density + ticks',
               family = binomial(link = 'logit'))
plot_df <- augment(fit_glm, type.predict = 'response')

plot_df %>%
  mutate(four = ifelse(directions == 4, density, NA),
         eight = ifelse(directions == 8, density, NA)) %>%
  ggplot(aes(x = density)) +
  geom_line(aes(x = density, y = .fitted,
                colour = burn_percentage)) +
  labs(x = "density", y = "directions") +
  geom_histogram(aes(x = four, y = stat(count)/1000), bins = 30, na.rm = TRUE) +
  geom_histogram(aes(x = eight, y = -1*stat(count/1000)), bins = 30, na.rm = TRUE, position = position_nudge(y = 1))
```
