---
title: "Fire_05_Calibration_Splines"
author: "Femke Keij S2647168"
date: "2023-02-28"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preliminaries
Clear working directory & set random seed:
```{r}
rm(list = ls(all = TRUE))

set.seed(42)
```

Packages used:
```{r}
# for importing / working with tidy data
library(tidyverse)

# for ggplot
library(directlabels)
library(patchwork)
library(ggbeeswarm)

# for fitting the GAM / splines
library(mgcv)
```

Source files for error computation and plotting functions:
```{r}
source('scripts/Fire model/Fire_ErrorFunctions.R')
```

Read in the data:
```{r}
# complete output data
fire_output <- read_csv('data/raw/fire_output.csv')

# training data
fire_train <- read_csv(
  'data/processed/fire_train.csv')
fire_train$directions <- as.factor(fire_train$directions)

# test data
fire_test <- read_csv(
  'data/processed/fire_test.csv')
fire_test$directions <- as.factor(fire_test$directions)
```

Random samples:
5, 10, 20, 30, 40, 50, 60, 70, and 80 $%$ of the total parameter space.
```{r}
N <- nrow(fire_output)
# range of (training) sample sizes to work with
n <- c(0.05 * N, 0.1 * N, 0.2 * N, 0.3 * N, 0.4 * N, 0.5 * N,
       0.6 * N, 0.7 * N, nrow(fire_train))
```

# Fitting
Here I'm fitting a cubic smoothing spline to predict the density and directions. In order to also include all variables, I incorporate everything into a GAM.
```{r}
GamFitting <- function(ticks = FALSE){
  # table to store all output
  length_tib <- sum(nrow(fire_test) * length(n))
  gam_results <- tibble(direction_true = numeric(length_tib),
                              direction_pred = numeric(length_tib),
                              density_true = numeric(length_tib),
                              density_pred = numeric(length_tib),
                              burn_true = numeric(length_tib),
                              burn_pred = numeric(length_tib),
                              n = numeric(length_tib),
                              ticks_included = numeric(length_tib))
  # vector to store mean density in each training sample
  # (to calculate point prediction performance later on)
  density_means <- numeric(length(n))
  # list to store fitted GAMs
  gam_fits <- list()

  count <- 1
  
  # for each sample size
  for(i in 1:length(n)){
    # sample training set
    ind_sample <- sample(1:nrow(fire_train), size = n[i])
    fire_train_sub <- fire_train[ind_sample, ]
    # calculate and save mean density in training set
    density_means[i] <- mean(fire_train_sub$density)
    
    # determine number of knots
    k_burn <- length(unique(round(fire_train_sub$burn_percentage)))
    
    if(ticks == TRUE){
      # GAM to predict density when ticks are included
      fit_gam_density <- gam(formula = density ~
                              s(burn_percentage, bs = 'cr', k = k_burn) +
                              factor(directions) + s(ticks),
             data = fire_train_sub)
      # GAM to predict directions when ticks are included
      fit_gam_directions <- gam(formula = directions ~ density +
                                  s(ticks) +
                                  s(burn_percentage, bs = 'cr', k = k_burn),
                                data = fire_train_sub, 
                                family = binomial(link = 'logit'))
      
    } else{
      # GAM to predict density when ticks are excluded
      fit_gam_density <- gam(formula = density ~
                              s(burn_percentage, bs = 'cr', k = k_burn) +
                              factor(directions),
             data = fire_train_sub)
      # GAM to predict directions when ticks are excluded
      fit_gam_directions <- gam(formula = directions ~ density +
                                  s(burn_percentage, bs = 'cr', k = k_burn),
                                data = fire_train_sub, 
                                family = binomial(link = 'logit'))
    }
    # predict density in test data using fitted GAM
    # include SEs to compute coverage later on
    predict_gam_density <- predict.gam(fit_gam_density,
                                       newdata = fire_test,
                                       se.fit = TRUE)
    # predict direction in test data using fitted regression
    predict_gam_directions <- predict.gam(fit_gam_directions,
                                          newdata = fire_test,
                                          type = 'response')
    # correct the predicted directions from '1' and '2' to '4' and '8'
    # to match test data
    for(k in 1:nrow(fire_test)){
      if(predict_gam_directions[k] < 0.5){
        predict_gam_directions[k] <- 4
      } else{
        predict_gam_directions[k] <- 8
      }
    }
    
    # for each instance of the test data
    for(j in 1:nrow(fire_test)){
      # find an instance in the training data that matches the directions
      # and density
      ind <- which(fire_train$density == round(predict_gam_density$fit[j]) &
                     fire_train$directions ==
                     predict_gam_directions[j])
      # if there are multiple, pick one
      if(length(ind) > 1){
        ind <- sample(ind, size = 1)
        # if there are none, increase density by 1 and see if there
        # is a matching instance in the training data
        # if not, continue until a matching instance is found
      } else if (length(ind) == 0){
        add <- 1
        while(length(ind) != 1){
          if(round(predict_gam_density$fit[j]) > 50){
            add <- add * -1
          }
          ind <- which(fire_train$density ==
                         (round(predict_gam_density$fit[j]) + add) &
                     fire_train$directions ==
                     predict_gam_directions[j])
          if(length(ind) > 1){
            ind <- sample(ind, size = 1)
          }
          add <- abs(add) + 1
        }
      }
      # store all results in table
      gam_results$direction_true[count] <- fire_test$directions[j]
      gam_results$direction_pred[count] <-
        predict_gam_directions[j]
      gam_results$density_true[count] <- fire_test$density[j]
      gam_results$density_pred[count] <- predict_gam_density$fit[j]
      gam_results$burn_true[count] <- fire_test$burn_percentage[j]
      gam_results$burn_pred[count] <- fire_train$burn_percentage[ind]
      gam_results$n[count] <- n[i]
      gam_results$ticks_included[count] <- if(ticks){'yes'}else{'no'}
      
      count <- count + 1
    }
    # store regression fit
    gam_fits <- append(gam_fits, fit_gam_density)
    gam_fits <- append(gam_fits, fit_gam_directions)
    
    print(paste('finished fitting sample size ', n[i]))
  }
  
  # correct format of direction_true variable
  gam_results %>%
    mutate(direction_true = replace(direction_true, direction_true == 1, 4),
         direction_true = replace(direction_true, direction_true == 2, 8)) ->
    gam_results
  
  # return list with results, density means, and fitted regression lines
  return(list(gam_results, density_means, gam_fits))
}
```

```{r}
# fit regressions with ticks
gam_results_ticks <- GamFitting(ticks = TRUE)
density_means_ticks <- gam_results_ticks[[2]]
gam_fits_ticks <- gam_results_ticks[[3]]
gam_results_ticks <- gam_results_ticks[[1]]

# fit regressions without ticks
gam_results_noticks <- GamFitting(ticks = FALSE)
density_means_noticks <- gam_results_noticks[[2]]
gam_fits_noticks <- gam_results_noticks[[3]]
gam_results_noticks <- gam_results_noticks[[1]]
```

```{r}
# run error calculations
gam_errors_ticks <- ComputeErrors(gam_results_ticks,
                                         ticks = TRUE,
                                         density_means_ticks,
                                         pred_int_included = FALSE)
gam_errors_noticks <- ComputeErrors(gam_results_noticks,
                                           ticks = FALSE,
                                           density_means_noticks,
                                           pred_int_included = FALSE)
```

Merge results
```{r}
gam_results <- rbind(gam_results_noticks,
                            gam_results_ticks)

gam_errors <- rbind(gam_errors_noticks, gam_errors_ticks)
``` 

# Plots
# Plots
## Errors
Fix results and errors data frames:
```{r}
gam_results <- gam_results %>%
  mutate(n = factor(n),
         direction_true = factor(direction_true),
         direction_pred = factor(direction_pred),
         direction_correct = (direction_pred == direction_true))

gam_errors <- gam_errors %>%
  mutate(n = factor(n))
```

Predicted vs. true density (largest sample size only):
```{r}
PlotPredTrueDensity(gam_results, n = 1584)

ggsave('figures/fire_gam_predtrue_density.pdf')
```
Predicted vs. true burn percentage (largest sample size only):
```{r}
PlotPredTrueBurn(gam_results, n = 1584)

ggsave('figures/fire_gam_predtrue_burn.pdf')
```
% correctly predicted parameters (density & directions):
```{r}
PlotPercCorrectParams(gam_errors, n_plot = c(99, 792, 1584))

ggsave('figures/fire_gam_perc_correct.pdf')
```
RMSE, NRMSE and point prediction performance for density parameter:
```{r}
PlotRMSEParams(gam_errors, n_plot = c(99, 792, 1584))

ggsave('figures/fire_gam_RMSE_density.pdf')
```
RMSE for burn percentage:
```{r}
PlotRMSEOut(gam_errors, n_plot = c(99, 792, 1584))

ggsave('figures/fire_gam_RMSE_burn.pdf')
```

## Model fit


