---
title: "Fire_04_Calibration_PiecewiseLinearRegression"
author: "Femke Keij S2647168"
date: "2023-02-22"
output: pdf_document
---
# Preliminaries
Clear working directory & set random seed:
```{r}
rm(list = ls(all = TRUE))

set.seed(42)
```

Packages used:
```{r}
# for piecewise linear regression
library(segmented)

# for importing / working with tidy data
library(tidyverse)

# for ggplot
library(directlabels)
library(patchwork)
library(ggbeeswarm)
```

Source files for error computation and plotting functions:
```{r}
source('scripts/Fire model/Fire_ErrorFunctions.R')
```

Read in the data:
```{r}
# complete output data
fire_output <- read_csv('data/raw/fire_output.csv')

# training data
fire_train <- read_csv(
  'data/processed/fire_train.csv')
fire_train$directions <- as.factor(fire_train$directions)

# test data
fire_test <- read_csv(
  'data/processed/fire_test.csv')
fire_test$directions <- as.factor(fire_test$directions)
```

Random samples:
5, 10, 20, 30, 40, 50, 60, 70, and 80 $%$ of the total parameter space.
```{r}
N <- nrow(fire_output)
# range of (training) sample sizes to work with
n <- c(0.05 * N, 0.1 * N, 0.2 * N, 0.3 * N, 0.4 * N, 0.5 * N,
       0.6 * N, 0.7 * N, nrow(fire_train))
```

# Fitting
Here, I'm fitting a piecewise linear regression to predict the density.
```{r}
PiecewiseRegressionFitting <- function(ticks = FALSE){
  # table to store all output
  length_tib <- sum(nrow(fire_test) * length(n))
  regression_results <- tibble(direction_true = numeric(length_tib),
                              direction_pred = numeric(length_tib),
                              density_true = numeric(length_tib),
                              density_pred = numeric(length_tib),
                              density_pred_lwr = numeric(length_tib),
                              density_pred_upr = numeric(length_tib),
                              burn_true = numeric(length_tib),
                              burn_pred = numeric(length_tib),
                              n = numeric(length_tib),
                              ticks_included = numeric(length_tib))
  # vector to store mean density in each training sample
  # (to calculate point prediction performance later on)
  density_means <- numeric(length(n))
  # list to store fitted regression lines
  regression_fits <- list()

  count <- 1
  
  # for each sample size
  for(i in 1:length(n)){
    # sample training set
    ind_sample <- sample(1:nrow(fire_train), size = n[i])
    fire_train_sub <- fire_train[ind_sample, ]
    # calculate and save mean density in training set
    density_means[i] <- mean(fire_train_sub$density)
    
    if(ticks == TRUE){
      # linear regression to predict density when ticks are included
      fit_lm_density <- lm(data = fire_train_sub,
                               formula = 'density ~ burn_percentage + directions + ticks')
      # logistic regression to predict directions when ticks are included
      fit_glm_directions <- glm(data = fire_train_sub,
                                    formula = 'directions ~ burn_percentage + density + ticks',
                                    family = binomial(link = 'logit'))
      
    } else{
      # linear regression to predict density when ticks are excluded
      fit_lm_density <- lm(data = fire_train_sub,
                               formula = 'density ~ burn_percentage + directions')
      # logistic regression to predict directions when ticks are excluded
      fit_glm_directions <- glm(data = fire_train_sub,
                                      formula = 'directions ~ burn_percentage + density',
                                      family = binomial(link = 'logit'))
    }
    # fit piecewise linear regression based on linear regression
    seg_fit_lm_density <- segmented(fit_lm_density, seg.Z = ~ burn_percentage,
                                    npsi = 2)
    # predict density in test data using fitted regression
    # include prediction interval to compute coverage later on
    predict_lm_density <- predict.segmented(seg_fit_lm_density,
                                             newdata = fire_test,
                                     interval = 'prediction')
    # predict direction in test data using fitted regression
    predict_glm_directions <- predict.glm(fit_glm_directions,
                                                  newdata = fire_test,
                                                  type = 'response')
    # correct the predicted directions from '1' and '2' to '4' and '8'
    # to match test data
    for(k in 1:nrow(fire_test)){
      if(predict_glm_directions[k] < 0.5){
        predict_glm_directions[k] <- 4
      } else{
        predict_glm_directions[k] <- 8
      }
    }
    
    # for each instance of the test data
    for(j in 1:nrow(fire_test)){
      # find an instance in the training data that matches the directions
      # and density
      ind <- which(fire_train$density == round(predict_lm_density[j,1]) &
                     fire_train$directions ==
                     predict_glm_directions[j])
      # if there are multiple, pick one
      if(length(ind) > 1){
        ind <- sample(ind, size = 1)
        # if there are none, increase density by 1 and see if there
        # is a matching instance in the training data
        # if not, continue until a matching instance is found
      } else if (length(ind) == 0){
        add <- 1
        while(length(ind) != 1){
          if(round(predict_lm_density[j,1]) > 50){
            add <- add * -1
          }
          ind <- which(fire_train$density ==
                         (round(predict_lm_density[j,1]) + add) &
                     fire_train$directions ==
                     predict_glm_directions[j])
          if(length(ind) > 1){
            ind <- sample(ind, size = 1)
          }
          add <- add + 1
        }
      }
      # find predicted burn percentage based on training data
      burn_pred <- fire_train$burn_percentage[ind]
      
      # store all results in table
      regression_results$direction_true[count] <- fire_test$directions[j]
      regression_results$direction_pred[count] <-
        predict_glm_directions[j]
      regression_results$density_true[count] <- fire_test$density[j]
      regression_results$density_pred[count] <- predict_lm_density[j,1]
      regression_results$density_pred_lwr[count] <-
        predict_lm_density[j,2]
      regression_results$density_pred_upr[count] <-
        predict_lm_density[j,3]
      regression_results$burn_true[count] <- fire_test$burn_percentage[j]
      regression_results$burn_pred[count] <- burn_pred
      regression_results$n[count] <- n[i]
      regression_results$ticks_included[count] <- if(ticks){'yes'}else{'no'}
      
      count <- count + 1
    }
    # store regression fit
    regression_fits <- append(regression_fits, fit_lm_density)
  }
  
  # correct format of direction_true variable
  regression_results %>%
    mutate(direction_true = replace(direction_true, direction_true == 1, 4),
         direction_true = replace(direction_true, direction_true == 2, 8)) ->
    regression_results
  
  # return table with results, density means, and fitted regression lines
  return(list(regression_results, density_means, regression_fits))
}
```

```{r}
# fit piecewise regressions with ticks
regression_results_ticks <- PiecewiseRegressionFitting(ticks = TRUE)
density_means_ticks <- regression_results_ticks[[2]]
regression_fits_ticks <- regression_results_ticks[[3]]
regression_results_ticks <- regression_results_ticks[[1]]

# fit piecewise regressions without ticks
regression_results_noticks <- PiecewiseRegressionFitting(ticks = FALSE)
density_means_noticks <- regression_results_noticks[[2]]
regression_fits_noticks <- regression_results_noticks[[3]]
regression_results_noticks <- regression_results_noticks[[1]]
```

```{r}
# run error calculations
regression_errors_ticks <- ComputeErrors(regression_results_ticks,
                                         ticks = TRUE,
                                         density_means_ticks,
                                         pred_int_included = TRUE)
regression_errors_noticks <- ComputeErrors(regression_results_noticks,
                                           ticks = FALSE,
                                           density_means_noticks,
                                           pred_int_included = TRUE)
```

Merge results
```{r}
regression_results <- rbind(regression_results_noticks,
                            regression_results_ticks)
write_csv(regression_results,
          'data/processed/fire_piecewise_results.csv')

regression_errors <- rbind(regression_errors_noticks, regression_errors_ticks)
write_csv(regression_errors,
          'data/processed/fire_piecewise_errors.csv')
``` 

# Plots
```{r}
# regression_results <- read_csv('data/processed/fire_piecewise_results.csv')

# regression_errors <- read_csv('data/processed/fire_piecewise_errors.csv')
```

## Errors
Fix results and errors data frames:
```{r}
regression_results <- regression_results %>%
  mutate(n = factor(n),
         direction_true = factor(direction_true),
         direction_pred = factor(direction_pred),
         direction_correct = (direction_pred == direction_true))

regression_errors <- regression_errors %>%
  mutate(n = factor(n))
```

Predicted vs. true density (largest sample size only):
```{r}
PlotPredTrueDensity(regression_results, n = 1584)

ggsave('figures/fire_piecewise_predtrue_density.pdf')
```
Predicted vs. true burn percentage (largest sample size only):
```{r}
PlotPredTrueBurn(regression_results, n = 1584)

ggsave('figures/fire_piecewise_predtrue_burn.pdf')
```
% correctly predicted parameters (density & directions):
```{r}
PlotPercCorrectParams(regression_errors, n_plot = c(99, 792, 1584))

ggsave('figures/fire_piecewise_perc_correct.pdf')
```
RMSE, NRMSE and point prediction performance for density parameter:
```{r}
PlotRMSEParams(regression_errors, n_plot = c(99, 792, 1584))

ggsave('figures/fire_piecewise_RMSE_density.pdf')
```
RMSE for burn percentage:
```{r}
PlotRMSEOut(regression_errors, n_plot = c(99, 792, 1584))

ggsave('figures/fire_piecewise_RMSE_burn.pdf')
```
Coverage:
```{r}
PlotCoverageDensity(regression_errors, n_plot = n)

ggsave('figures/fire_piecewise_coverage.pdf')
```

## Model fit
Piecewise linear regression for density (only for largest training sample size)
```{r}
fit_ticks <- lm(data = fire_train,
          formula = density ~ directions + ticks + burn_percentage)
segmented_fit_ticks <- segmented(fit_ticks,
                                 seg.Z = ~ burn_percentage, npsi = 2)

fit_noticks <- lm(data = fire_train,
          formula = density ~ directions + burn_percentage)
segmented_fit_noticks <- segmented(fit_noticks,
                                   seg.Z = ~ burn_percentage, npsi = 2)

dat_ticks <- data.frame(burn_percentage = fire_train$burn_percentage,
                        density = fire_train$density,
                        fit = broken.line(segmented_fit_ticks)$fit,
                        directions = fire_train$directions)
dat_noticks <- data.frame(burn_percentage = fire_train$burn_percentage,
                          density = fire_train$density,
                          fit = broken.line(segmented_fit_noticks)$fit,
                          directions = fire_train$directions)
row.names(dat_ticks) <- NULL
row.names(dat_noticks) <- NULL

p1 <- ggplot(data = fire_train, aes(x = burn_percentage, y = density,
                              colour = directions,
                              group = directions)) +
  geom_point() +
  geom_line(data = dat_noticks, aes(x = burn_percentage, y = fit,
                                    group = directions,
                                    colour = directions)) +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
        panel.border = element_rect(colour = 'lightgrey', fill = NA)) +
  labs(x = '% burned trees at last tick',
       y = 'tree density (%)') +
  ggtitle('no ticks') +
  coord_flip()

p2 <- ggplot(data = fire_train, aes(x = burn_percentage, y = density,
                              colour = directions)) +
  geom_point() +
  geom_line(data = dat_ticks, aes(x = burn_percentage, y = fit,
                                  group = directions)) +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
        panel.border = element_rect(colour = 'lightgrey', fill = NA)) +
  labs(x = '% burned trees at last tick',
       y = 'tree density (%)') +
  ggtitle('ticks') +
  coord_flip()

patch <- p1 + p2 +
    plot_layout(guides = 'collect') +
    plot_annotation(tag_levels = 'A') & 
    theme(plot.tag = element_text(size = 8)) &
    ylab(NULL)
  
wrap_elements(panel = patch) +
  labs(tag = 'tree density(%)') +
  theme(plot.tag = element_text(size = rel(1)),
        plot.tag.position = 'bottom')
```