---
title: "Fire_09_MultivariateRandomForest"
author: "Femke Keij S2647168"
date: "2023-03-21"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preliminaries
Clear working directory & set random seed:
```{r}
rm(list = ls(all = TRUE))

set.seed(42)
```

Packages used:
```{r}
# for importing / working with tidy data
library(tidyverse)

# for ggplot
library(directlabels)
library(patchwork)
library(ggbeeswarm)

# for cross-validation folds
library(caret)

# for random forest
library(randomForest)

# for multivariate random forest
library(MultivariateRandomForest)
```

Source files for error computation and plotting functions:
```{r}
source('scripts/Fire model/Fire_Functions.R')
```

# Fitting
A function to fit a Partial Least Squares for the density and direction parameters, using 5-fold cross-validation.
```{r}
FireRFFitting <- function(data_clean,
                       data_noise){
  # data_clean: the data frame containing the input parameters as well as the relevant outputs
  # data_noise: supply data with noise here
  
  # create the folds
  flds <- createFolds(1:nrow(data_clean),
                      k = 5, list = TRUE, returnTrain = FALSE)

  # to store the results
  results_df <- tibble(density = numeric(),
                       directions = numeric(),
                       density_pred = numeric(),
                       directions_pred = numeric(),
                       noise = character(),
                       fold = numeric())
  
  # for each cross-validation fold
  for(i in 1:5){
    # remove the test fold to create the training data
    dat_train <- data_clean %>%
      filter(!row_number() %in% unlist(flds[i]))
    
    # obtain the test fold without noise
    dat_test_clean <- data_clean %>%
        filter(row_number() %in% unlist(flds[i]))
    # obtain the test fold with noise
    dat_test_noise <- data_noise %>%
        filter(row_number() %in% unlist(flds[i]))
    
    # train the random forest
    fit_rf_density <- randomForest(density ~ . - directions,
                                      data = dat_train,
                                   ntree = 500)
    fit_rf_directions <- randomForest(as.factor(directions) ~
                                        . - density,
                                      data = dat_train,
                                      proximity = TRUE,
                                      ntree = 500)
    
    # predict the density & directions in the left-out fold without noise
    predict_rf_clean_density <- predict(fit_rf_density,
                                        newdata = dat_test_clean)
    predict_rf_clean_directions <-
      as.numeric(as.character(predict(fit_rf_directions,
                         newdata = dat_test_clean)))
    
    # predict the density & directions in the left-out fold with noise
    predict_rf_noise_density <- predict(fit_rf_density,
                                        newdata = dat_test_noise)
    predict_rf_noise_directions <- 
      as.numeric(as.character(predict(fit_rf_directions,
                                      newdata = dat_test_noise)))
    
    # add the predictions to the data frame
    results_new_clean <- tibble(density = dat_test_clean$density,
                          directions = dat_test_clean$directions,
                          density_pred = 
                            predict_rf_clean_density,
                          directions_pred = 
                            predict_rf_clean_directions,
                          fold = i,
                          noise = 'clean')
    results_new_noise <- tibble(density = dat_test_clean$density,
                                directions = dat_test_clean$directions,
                                density_pred =
                                  predict_rf_noise_density,
                                directions_pred = 
                                  predict_rf_noise_directions,
                                fold = i,
                                noise = 'noise')
    results_df <- results_df %>%
      add_row(results_new_clean) %>%
      add_row(results_new_noise)
  }
  
  return(results_df)
}
```

Timesteps, not summarised
```{r}
data_time_full_clean <-
  read_csv('data/training/fire_time_full_clean.csv')
data_time_full_outputnoise <-
  read_csv('data/training/fire_time_full_outputnoise.csv')

# fit regressions
fit_data_time_full <-
  FireRFFitting(data_time_full_clean,
                 data_noise = data_time_full_outputnoise)

# compute errors
errors_data_time_full <- 
  FireComputeErrors(fit_data_time_full,
                    sample_size = 198,
                    sample_method = 'full',
                    summarise_runs = 'full',
                    datapoints = 'timepoints')

# dataframe to store all errors
errors <- errors_data_time_full
```

Timesteps, summarised
```{r}
data_time_sum_clean <-
  read_csv('data/training/fire_time_sum_clean.csv')
data_time_sum_outputnoise <-
  read_csv('data/training/fire_time_sum_outputnoise.csv')

fit_data_time_sum <-
  FireRFFitting(data_time_sum_clean,
                 data_noise = data_time_sum_outputnoise)

errors_data_time_sum <- 
  FireComputeErrors(fit_data_time_sum,
                    sample_size = 198,
                    sample_method = 'full',
                    summarise_runs = 'summarised',
                    datapoints = 'timepoints')

errors <- errors %>%
  add_row(errors_data_time_sum)
```

Endpoints, not summarised
```{r}
data_end_full_clean <-
  read_csv('data/training/fire_end_full_clean.csv')
data_end_full_outputnoise <-
  read_csv('data/training/fire_end_full_outputnoise.csv')

fit_data_end_full <-
  FireRFFitting(data_end_full_clean,
                 data_noise = data_end_full_outputnoise)

errors_data_end_full <- 
  FireComputeErrors(fit_data_end_full,
                    sample_size = 198,
                    sample_method = 'full',
                    summarise_runs = 'full',
                    datapoints = 'endpoints')

errors <- errors %>%
  add_row(errors_data_end_full)
```

Endpoints, summarised
```{r}
data_end_sum_clean <- read_csv('data/training/fire_end_sum_clean.csv')
data_end_sum_outputnoise <-
  read_csv('data/training/fire_end_sum_outputnoise.csv')

fit_data_end_sum <-
  FireRFFitting(data_end_sum_clean,
                 data_noise = data_end_sum_outputnoise)

errors_data_end_sum <- 
  FireComputeErrors(fit_data_end_sum,
                    sample_size = 198,
                    sample_method = 'full',
                    summarise_runs = 'summarised',
                    datapoints = 'endpoints')

errors <- errors %>%
  add_row(errors_data_end_sum)
```

Save results:
```{r}
errors %>%
  mutate(algorithm = 'Random Forest') %>%
  write_csv(file = 'data/results/fire_rf_errors.csv')
```

# Plots
```{r}
plot1 <- PlotPercCorrectParams(errors)

ggsave('figures/fire_rf_perc_correct.pdf')
```

```{r}
plot2 <- PlotErrorsDensity(errors)

ggsave('figures/fire_rf_density.pdf')
```

```{r}
plot3 <- PlotErrorsDirections(errors)

ggsave('figures/fire_rf_directions.pdf')
```