---
title: "Fire_03_Calibration_SimpleRegression"
author: "Femke Keij S2647168"
date: "2023-02-14"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preliminaries
Clear working directory & set random seed:
```{r}
rm(list = ls(all = TRUE))

set.seed(42)
```

Packages used:
```{r}
# for importing / working with tidy data
library(tidyverse)

# for creating cross-validation folds
library(caret)

# for error computation
library(mltools)

# for ggplot
library(directlabels)
library(patchwork)
library(ggbeeswarm)
```

Functions to compute and plot errors:
```{r}
source('scripts/Fire model/Fire_Functions.R')
```

# Fitting
A function to fit a simple linear regression for the density parameter and a logistic regression for the directions parameter, using 5-fold cross-validation.
```{r}
FireSimpleRegressionFitting <- function(data_clean,
                                        data_noise){
  # data_clean: the data frame containing the input parameters as well as the relevant outputs
  # data_noise: supply data with noise here
  
  # create the folds
  flds <- createFolds(1:nrow(data_clean),
                      k = 5, list = TRUE, returnTrain = FALSE)

  # to store the results
  results_df <- tibble(density = numeric(),
                       directions = numeric(),
                       density_pred = numeric(),
                       directions_pred = numeric(),
                       noise = character(),
                       fold = numeric())
  
  # set directions 4 to 0 and 8 to 1 (eases interpretation of logistic regression)
  data_clean <- data_clean %>%
    mutate(directions = ifelse(directions == 4, 0, 1))
  data_noise <- data_noise %>%
      mutate(directions = ifelse(directions == 4, 0, 1))
  
  # for each cross-validation fold
  for(i in 1:5){
    # remove the test fold to create the training data
    dat_train <- data_clean %>%
      filter(!row_number() %in% unlist(flds[i]))
    
    # obtain the test fold without noise
    dat_test_clean <- data_clean %>%
        filter(row_number() %in% unlist(flds[i]))
    # obtain the test fold with noise
    dat_test_noise <- data_noise %>%
        filter(row_number() %in% unlist(flds[i]))
    
    # train the linear & logistic regressions
    fit_lm_density <- lm(formula = density ~ . - directions,
                         data = dat_train)
    fit_glm_directions <- glm(formula = directions ~ . - density,
                              data = dat_train)
    
    # predict the density & directions in the left-out fold without noie
    predict_density_clean <- predict.lm(fit_lm_density,
                                  newdata = dat_test_clean,
                                  interval = 'prediction')
    predict_directions_clean <- predict.glm(fit_glm_directions,
                                      newdata = dat_test_clean,
                                      type = 'response')
    # predict the density & directions in the left-out fold with noise
    predict_density_noise <- predict.lm(fit_lm_density,
                                  newdata = dat_test_noise,
                                  interval = 'prediction')
    predict_directions_noise <- predict.glm(fit_glm_directions,
                                      newdata = dat_test_noise,
                                      type = 'response')
    
    
    # correct predictions for directions to 4 vs. 8
    for(j in 1:length(predict_directions_clean)){
      predict_directions_clean[j] <- ifelse(
        predict_directions_clean[j] <= 0.5,
                                      4, 8)
    }
    for(j in 1:length(predict_directions_noise)){
      predict_directions_noise[j] <- ifelse(
        predict_directions_noise[j] <= 0.5,
                                      4, 8)
    }
    
    # add the predictions to the data frame
    results_new_clean <- tibble(density = dat_test_clean$density,
                          directions = dat_test_clean$directions,
                          density_pred = 
                            predict_density_clean[, 1],
                          directions_pred =
                            predict_directions_clean,
                          fold = i,
                          noise = 'clean')
    results_new_noise <- tibble(density = dat_test_clean$density,
                                directions = dat_test_clean$directions,
                                density_pred =
                                  predict_density_noise[, 1],
                                directions_pred = 
                                  predict_directions_noise,
                                fold = i,
                                noise = 'noise')
    results_df <- results_df %>%
      add_row(results_new_clean) %>%
      add_row(results_new_noise)
  }
  # correct back to 4 and 8 directions
  results_df <- results_df %>%
    mutate(directions = ifelse(directions == 0, 4, 8))
  
  return(results_df)
}
```

Now run these functions for each data combination. In this case:
- endpoints only, time steps version here: endpoint, middle point, total number of ticks, mean, min, max, sd, trend
- runs summarised or not
- data without noise, data with output noise (no temporal noise for this ABM)
There are no variations with different sample sizes or sampling methods here, since I sampled the full parameter space for this model. The 10, 20, and 50 timesteps are excluded due to collinearity issues.

Timesteps, not summarised
```{r}
data_time_full_clean <-
  read_csv('data/training/fire_time_full_clean.csv')
data_time_full_outputnoise <-
  read_csv('data/training/fire_time_full_outputnoise.csv')

# fit regressions
fit_data_time_full <-
  FireSimpleRegressionFitting(data_time_full_clean,
                              data_noise = data_time_full_outputnoise)

# compute errors
errors_data_time_full <- 
  FireComputeErrors(fit_data_time_full,
                    sample_size = 198,
                    sample_method = 'full',
                    summarise_runs = 'full',
                    datapoints = 'timepoints')

# dataframe to store all errors
errors <- errors_data_time_full
```

Timesteps, summarised
```{r}
data_time_sum_clean <-
  read_csv('data/training/fire_time_sum_clean.csv')
data_time_sum_outputnoise <-
  read_csv('data/training/fire_time_sum_outputnoise.csv')

fit_data_time_sum <-
  FireSimpleRegressionFitting(data_time_sum_clean,
                              data_noise = data_time_sum_outputnoise)
errors_data_time_sum <- 
  FireComputeErrors(fit_data_time_sum,
                    sample_size = 198,
                    sample_method = 'full',
                    summarise_runs = 'summarised',
                    datapoints = 'timepoints')

errors <- errors %>%
  add_row(errors_data_time_sum)
```

Endpoints, not summarised
```{r}
data_end_full_clean <-
  read_csv('data/training/fire_end_full_clean.csv')
data_end_full_outputnoise <-
  read_csv('data/training/fire_end_full_outputnoise.csv')

fit_data_end_full <-
  FireSimpleRegressionFitting(data_end_full_clean,
                              data_noise = data_end_full_outputnoise)

errors_data_end_full <- 
  FireComputeErrors(fit_data_end_full,
                    sample_size = 198,
                    sample_method = 'full',
                    summarise_runs = 'full',
                    datapoints = 'endpoints')

errors <- errors %>%
  add_row(errors_data_end_full)
```

Endpoints, summarised
```{r}
data_end_sum_clean <- read_csv('data/training/fire_end_sum_clean.csv')
data_end_sum_outputnoise <-
  read_csv('data/training/fire_end_sum_outputnoise.csv')

fit_data_end_sum <-
  FireSimpleRegressionFitting(data_end_sum_clean,
                              data_noise = data_end_sum_outputnoise)
errors_data_end_sum <- 
  FireComputeErrors(fit_data_end_sum,
                    sample_size = 198,
                    sample_method = 'full',
                    summarise_runs = 'summarised',
                    datapoints = 'endpoints')

errors <- errors %>%
  add_row(errors_data_end_sum)
```

Save results:
```{r}
errors %>%
  mutate(algorithm = 'Linear Regression') %>%
  write_csv(file = 'data/results/fire_lr_errors.csv')
```

# Plots
% correctly predicted parameters (density & directions):
```{r}
plot1 <- PlotPercCorrectParams(errors)

ggsave('figures/fire_regression_perc_correct.pdf')
```
RMSE, NRMSE, point prediction performance, and coverage for density parameter:
```{r}
plot2 <- PlotErrorsDensity(errors)

ggsave('figures/fire_regression_density_metrics.pdf')
```
F1, kappa score, and MCC for directions:
```{r}
plot3 <- PlotErrorsDirections(errors)

ggsave('figures/fire_regression_directions_metrics.pdf')
```

############### ------------- OLD CODE ------------------ #############
## Errors
Fix results and errors data frames:
```{r, eval = FALSE}
regression_results <- regression_results %>%
  mutate(n = factor(n),
         direction_true = factor(direction_true),
         direction_pred = factor(direction_pred),
         direction_correct = (direction_pred == direction_true))

regression_errors <- regression_errors %>%
  mutate(n = factor(n))
```
### Predicted vs. true
Predicted vs. true for density, directions, and burn percentage
```{r, eval = FALSE}
p1 <- PlotPredTrueDensity(regression_results, n = 1584)
p2 <- PlotPredTrueDirection(regression_results, n = 1584)
p3 <- PlotPredTrueBurn(regression_results, n = 1584)

p1 / p2 / p3
```
Predicted vs. true density (largest sample size only):
```{r, eval = FALSE}
PlotPredTrueDensity(regression_results, n = 1584)

ggsave('figures/fire_regression_predtrue_density.pdf')
```
Confusion matrix for predicted vs. true directions:
```{r, eval = FALSE}
PlotPredTrueDirection(regression_results, n = 1584)

ggsave('figures/fire_regression_predtrue_direction.pdf')
```
Predicted vs. true burn percentage (largest sample size only):
```{r, eval = FALSE}
PlotPredTrueBurn(regression_results, n = 1584)

ggsave('figures/fire_regression_predtrue_burn.pdf')
```
### Parameters
% correctly predicted parameters (density & directions):
```{r, eval = FALSE}
PlotPercCorrectParams(regression_errors, n_plot = c(99, 792, 1584))

ggsave('figures/fire_regression_perc_correct.pdf')
```
RMSE, NRMSE, point prediction performance, and coverage for density parameter:
```{r, eval = FALSE}
PlotErrorsDensity(regression_errors, n_plot = c(99, 792, 1584))

ggsave('figures/fire_regression_density_metrics.pdf')
```
F1, kappa score, and MCC for directions:
```{r, eval = FALSE}
PlotErrorsDirections(regression_errors, n_plot = c(99, 792, 1584))

ggsave('figures/fire_regression_directions_metrics.pdf')
```
### Outcome variable
RMSE for burn percentage:
```{r, eval = FALSE}
PlotRMSEOut(regression_errors, n_plot = c(99, 792, 1584))

ggsave('figures/fire_regression_RMSE_burn.pdf')
```

## Model fit
Fitted regression lines (smallest and largest sample size):
Linear regression for density
```{r, eval = FALSE}
RegressionLinePlot <- function(sample = 1, ticks = TRUE){
  if(ticks){
    formula <- 'density ~ burn_percentage + directions + ticks'
    subtitle <- 'with ticks'
  } else{
    formula <- 'density ~ burn_percentage + directions'
    subtitle <- 'without ticks'
  }
  
  subtitle <- paste(subtitle, ' sample size ', n[sample])
  
  ind_sample <- sample(1:nrow(fire_train), size = n[sample])
  fire_train_sub <- fire_train[ind_sample, ]
  
  fire_train_sub %>%
    ggplot(mapping = aes(x = burn_percentage, y = density,
                         colour = directions)) +
    geom_point() +
    geom_line(data = fortify(lm(data = fire_train_sub,
                                formula = formula)),
              aes(x = burn_percentage, y = .fitted)) +
    labs(x = '% burned trees at last tick',
         y = 'tree density (%)',
         subtitle = subtitle) +
    theme_minimal() +
    coord_flip() -> returnplot

  return(returnplot)
}

p1 <- RegressionLinePlot(sample = 1, ticks = TRUE)
p2 <- RegressionLinePlot(sample = 1, ticks = FALSE)
p3 <- RegressionLinePlot(sample = 9, ticks = TRUE)
p4 <- RegressionLinePlot(sample = 9, ticks = FALSE)

p1 + p2 + p3 + p4 +
  plot_layout(guides = 'collect') +
  plot_annotation(tag_levels = 'A') & 
  theme(plot.tag = element_text(size = 8))

ggsave('figures/fire_regression_fit.pdf')
```

Logistic regression for directions (only largest training sample size)
```{r, eval = FALSE}
fit_glm_ticks <- glm(data = fire_train,
               formula = 'directions ~ burn_percentage + density + ticks',
               family = binomial(link = 'logit'))
plot_df_ticks <- augment(fit_glm_ticks, type.predict = 'response')

p1 <- plot_df_ticks %>%
  ggplot(mapping = aes(x = density, y = .fitted,
                       colour = directions)) +
  geom_point(size = 0.5) +
  labs(x = 'tree density',
       y = 'probability number of directions is 8',
       colour = 'true number \n of directions') +
  geom_hline(yintercept = 0.5,
             linetype = 'dashed',
             colour = 'darkgreen') +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
        panel.border = element_rect(colour = 'lightgrey', fill = NA)) +
  annotate(geom = 'text',
           x = 17, y = 0.25,
           label = 'directions \n classified \n as 4',
           size = 3) +
  annotate(geom = 'text',
           x = 83, y = 0.75,
           label = 'directions \n classified \n as 8',
           size = 3) +
  ylim(c(0, 1)) +
  ggtitle('ticks')

fit_glm_noticks <- glm(data = fire_train,
               formula = 'directions ~ burn_percentage + density',
               family = binomial(link = 'logit'))
plot_df_noticks <- augment(fit_glm_noticks, type.predict = 'response')

p2 <- plot_df_noticks %>%
  ggplot(mapping = aes(x = density, y = .fitted,
                       colour = directions)) +
  geom_point(size = 0.5) +
  labs(x = 'tree density',
       y = 'probability number of directions is 8',
       colour = 'true number \n of directions') +
  geom_hline(yintercept = 0.5,
             linetype = 'dashed',
             colour = 'darkgreen') +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
        panel.border = element_rect(colour = 'lightgrey', fill = NA)) +
  annotate(geom = 'text',
           x = 17, y = 0.25,
           label = 'directions \n classified \n as 4',
           size = 3) +
  annotate(geom = 'text',
           x = 83, y = 0.75,
           label = 'directions \n classified \n as 8',
           size = 3) +
  ylim(c(0, 1)) +
  ggtitle('no ticks')

patch <- p1 + p2 +
    plot_layout(guides = 'collect') +
    plot_annotation(tag_levels = 'A') & 
    theme(plot.tag = element_text(size = 8)) &
    xlab(NULL)
  
wrap_elements(panel = patch) +
  labs(tag = 'tree density (%)') +
  theme(plot.tag = element_text(size = rel(1)),
        plot.tag.position = 'bottom')

ggsave('figures/fire_logistic_fit.pdf')
```