---
title: "Fire_06_Calibration_ElasticNet"
author: "Femke Keij S2647168"
date: "2023-02-28"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Preliminaries
Clear working directory & set random seed:
```{r}
rm(list = ls(all = TRUE))

set.seed(42)
```

Packages used:
```{r}
# for importing / working with tidy data
library(tidyverse)

# for ggplot
library(directlabels)
library(patchwork)
library(ggbeeswarm)

# for fitting the elastic net
library(caret)
library(glmnet)
```

Read in the data:
```{r}
# complete output data
fire_output <- read_csv('data/raw/fire_output.csv')

# training data
fire_train <- read_csv(
  'data/processed/fire_train.csv')
fire_train$directions <- as.factor(fire_train$directions)

# test data
fire_test <- read_csv(
  'data/processed/fire_test.csv')
fire_test$directions <- as.factor(fire_test$directions)
```

Random samples:
5, 10, 20, 30, 40, 50, 60, 70, and 80 $%$ of the total parameter space.
```{r}
N <- nrow(fire_output)
# range of (training) sample sizes to work with
n <- c(0.05 * N, 0.1 * N, 0.2 * N, 0.3 * N, 0.4 * N, 0.5 * N,
       0.6 * N, 0.7 * N, nrow(fire_train))
```

# Fitting
Here I'm fitting an elastic net regression to predict density and direction.
```{r}
fit_density <- train(density ~ burn_percentage + directions + ticks,
                     data = fire_train,
                     method = 'glmnet',
                     trControl = trainControl(method = 'cv', number = 10),
                     tuneLength = 10)

fit_direction <- train(directions ~ burn_percentage + density + ticks,
                       data = fire_train,
                       method = 'glmnet',
                       trControl = trainControl(method = 'cv', number = 10),
                       tuneLength = 10)

pred_density <- predict(fit_density, newdata = fire_test)
pred_direction <- predict(fit_direction, newdata = fire_test)

```


```{r}
ElasticNetFitting <- function(ticks = FALSE){
  # table to store all output
  length_tib <- sum(nrow(fire_test) * length(n))
  elastic_results <- tibble(direction_true = numeric(length_tib),
                              direction_pred = numeric(length_tib),
                              density_true = numeric(length_tib),
                              density_pred = numeric(length_tib),
                              burn_true = numeric(length_tib),
                              burn_pred = numeric(length_tib),
                              n = numeric(length_tib),
                              ticks_included = numeric(length_tib))
  # vector to store mean density in each training sample
  # (to calculate point prediction performance later on)
  density_means <- numeric(length(n))
  # list to store fitted Elastic Nets
  elastic_fits <- list()

  count <- 1
  
  # for each sample size
  for(i in 1:length(n)){
    # sample training set
    ind_sample <- sample(1:nrow(fire_train), size = n[i])
    fire_train_sub <- fire_train[ind_sample, ]
    # calculate and save mean density in training set
    density_means[i] <- mean(fire_train_sub$density)
    
    # determine number of knots
    k_burn <- length(unique(round(fire_train_sub$burn_percentage)))
    
    if(ticks == TRUE){
      # Elastic Net to predict density when ticks are included
      fit_elastic_density <- train(density ~ burn_percentage +
                                     directions + ticks,
                                   data = fire_train_sub,
                                   method = 'glmnet',
                                   trControl = 
                                     trainControl(method = 'cv', number = 10),
                                   tuneLength = 10)
      # Elastic Net to predict directions when ticks are included
      fit_elastic_directions <- train(directions ~ burn_percentage +
                                        density + ticks,
                                      data = fire_train_sub,
                                      method = 'glmnet',
                                      trControl =
                                        trainControl(method = 'cv',
                                                     number = 10),
                                      tuneLength = 10)
      
    } else{
      # Elastic Net to predict density when ticks are excluded
      fit_elastic_density <- train(density ~ burn_percentage +
                                     directions,
                                   data = fire_train_sub,
                                   method = 'glmnet',
                                   trControl = 
                                     trainControl(method = 'cv', number = 10),
                                   tuneLength = 10)
      # Elastic Net to predict directions when ticks are excluded
      fit_elastic_directions <- train(directions ~ burn_percentage +
                                        density,
                                      data = fire_train_sub,
                                      method = 'glmnet',
                                      trControl =
                                        trainControl(method = 'cv',
                                                     number = 10),
                                      tuneLength = 10)
    }
    # predict density in test data using fitted Elastic Net
    # include SEs to compute coverage later on
    predict_elastic_density <- predict(fit_elastic_density,
                                       newdata = fire_test)
    # predict direction in test data using fitted regression
    predict_elastic_directions <- predict(fit_elastic_directions,
                                          newdata = fire_test)
    
    # for each instance of the test data
    for(j in 1:nrow(fire_test)){
      # find an instance in the training data that matches the directions
      # and density
      ind <- which(fire_train$density == round(predict_elastic_density[j]) &
                     fire_train$directions ==
                     predict_elastic_directions[j])
      # if there are multiple, pick one
      if(length(ind) > 1){
        ind <- sample(ind, size = 1)
        # if there are none, increase density by 1 and see if there
        # is a matching instance in the training data
        # if not, continue until a matching instance is found
      } else if (length(ind) == 0){
        add <- 1
        while(length(ind) != 1){
          if(round(predict_elastic_density[j]) > 50){
            add <- add * -1
          }
          ind <- which(fire_train$density ==
                         (round(predict_elastic_density[j]) + add) &
                     fire_train$directions ==
                     predict_elastic_directions[j])
          if(length(ind) > 1){
            ind <- sample(ind, size = 1)
          }
          add <- abs(add) + 1
        }
      }
      # store all results in table
      elastic_results$direction_true[count] <- fire_test$directions[j]
      elastic_results$direction_pred[count] <-
        predict_elastic_directions[j]
      elastic_results$density_true[count] <- fire_test$density[j]
      elastic_results$density_pred[count] <- predict_elastic_density[j]
      elastic_results$burn_true[count] <- fire_test$burn_percentage[j]
      elastic_results$burn_pred[count] <- fire_train$burn_percentage[ind]
      elastic_results$n[count] <- n[i]
      elastic_results$ticks_included[count] <- if(ticks){'yes'}else{'no'}
      
      count <- count + 1
    }
    # store regression fit
    elastic_fits <- append(elastic_fits, fit_elastic_density)
    elastic_fits <- append(elastic_fits, fit_elastic_directions)
    
    print(paste('finished fitting sample size ', n[i]))
  }
  
  # correct format of direction_true variable
  elastic_results %>%
    mutate(direction_true = replace(direction_true, direction_true == 1, 4),
         direction_true = replace(direction_true, direction_true == 2, 8)) ->
    elastic_results
  
  # return list with results, density means, and fitted regression lines
  return(list(elastic_results, density_means, elastic_fits))
}
```

```{r}
# fit regressions with ticks
elastic_results_ticks <- ElasticNetFitting(ticks = TRUE)
density_means_ticks <- elastic_results_ticks[[2]]
elastic_fits_ticks <- elastic_results_ticks[[3]]
elastic_results_ticks <- elastic_results_ticks[[1]]

# fit regressions without ticks
elastic_results_noticks <- ElasticNetFitting(ticks = FALSE)
density_means_noticks <- elastic_results_noticks[[2]]
elastic_fits_noticks <- elastic_results_noticks[[3]]
elastic_results_noticks <- elastic_results_noticks[[1]]
```

Function to calculate the error measures for the regression lines.
The following function summarises the following:
- % of cases in which both direction and density are estimated correctly
- % of cases in which density is estimated correctly
- % of cases in which direction is estimated correctly
- % of cases in which the density is estimated to be within 10% of the true density
- RMSE of the estimated density
- NRMSE of the estimated density
- point prediction performance of the estimated density
```{r}
ElasticNetErrors <- function(regression_results, ticks = FALSE,
                             density_means){
  
  merge(regression_results, as_tibble(cbind(n, density_means)),
        by = c('n')) %>%
    group_by(n) %>%
    summarise(perc_correct = sum(direction_true ==
                                   direction_pred &
                                   density_true == round(density_pred))
                / nrow(fire_test),
              perc_density_correct = sum(density_true ==
                                         round(density_pred)) /
                nrow(fire_test),
              perc_direction_correct = sum(direction_true ==
                                           direction_pred) /
                nrow(fire_test),
              perc_correct_cat = sum(density_pred >=
                                       density_true - 5 &
                                       density_pred <=
                                       density_true + 5) /
                nrow(fire_test),
              RMSE = sqrt(sum((density_pred - density_true)^2))
                /nrow(fire_test),
              NRMSE = (sqrt(sum((density_pred -
                                   density_true)^2))
                     / nrow(fire_test)) / sd(density_true),
              point_pred_performance = 1 - 
                sum(sqrt((density_pred - density_true)^2)) /
                sum(sqrt((density_true - density_means)^2))) ->
    regression_errors
  
  regression_errors$ticks_included <- 
    if(ticks == TRUE){'yes'} else{'no'}

  return(regression_errors)
}
```

```{r}
# run error calculations
elastic_errors_ticks <- ElasticNetErrors(elastic_results_ticks,
                                            ticks = TRUE,
                                            density_means_ticks)
elastic_errors_noticks <- ElasticNetErrors(elastic_results_noticks,
                                            ticks = FALSE,
                                            density_means_noticks)
```

Merge results
```{r}
elastic_results <- rbind(elastic_results_noticks,
                            elastic_results_ticks)

elastic_errors <- rbind(elastic_errors_noticks, elastic_errors_ticks)
``` 

# Plots
Predicted vs. true density (largest sample size only):
```{r}
elastic_results$n <- as.factor(elastic_results$n)
elastic_results$direction_true <-
  as.factor(elastic_results$direction_true)
elastic_results$direction_pred <-
  as.factor(elastic_results$direction_pred)

supp_labs <- c('without ticks', 'with ticks')
names(supp_labs) <- c('no', 'yes')

elastic_results %>%
  filter(n == 1584) %>%
  ggplot(mapping = aes(x = density_pred, y = density_true,
                       colour = direction_true,
                       shape = direction_pred)) +
  facet_grid(. ~ ticks_included,
             labeller = labeller(ticks_included = supp_labs)) +
  geom_point() +
  theme_minimal() +
  labs(x = 'predicted density',
       y = 'true density',
       colour = 'True number of directions',
       shape = 'Predicted number of directions') +
  theme(panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank())

ggsave('figures/fire_ElasticNet_pred_vs_true.pdf')
```

% of predictions that are correct
```{r}
elastic_errors$n <- as.factor(elastic_errors$n)

elastic_errors %>%
  filter(n == c(99, 792, 1584)) %>%
  ggplot(mapping = aes(x = n, y = perc_correct,
                       fill = ticks_included)) +
  geom_bar(stat = 'identity',
           position = position_dodge()) +
  labs(y = 'density & direction',
       fill = 'Ticks included') +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()) -> p1

elastic_errors %>%
  filter(n == c(99, 792, 1584)) %>%
  ggplot(mapping = aes(x = n, y = perc_density_correct,
                       fill = ticks_included)) +
  geom_bar(stat = 'identity',
           position = position_dodge()) +
  labs(y = 'density',
       fill = 'Ticks included') +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()) -> p2

elastic_errors %>%
  filter(n == c(99, 792, 1584)) %>%
  ggplot(mapping = aes(x = n, y = perc_direction_correct,
                       fill = ticks_included)) +
  geom_bar(stat = 'identity',
           position = position_dodge()) +
  labs(y = 'direction',
       fill = 'Ticks included') +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()) -> p3

elastic_errors %>%
  filter(n == c(99, 792, 1584)) %>%
  ggplot(mapping = aes(x = n, y = perc_correct_cat,
                       fill = ticks_included)) +
  geom_bar(stat = 'identity',
           position = position_dodge()) +
  labs(y = 'density within \n 10% of true density',
       fill = 'Ticks included') +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()) -> p4

p1 + p2 + p3 + p4 +
  plot_layout(guides = 'collect') +
  plot_annotation(tag_levels = 'A',
                  title = '% correctly predicted') & 
  theme(plot.tag = element_text(size = 8))

ggsave('figures/fire_ElasticNet_perc_correct.pdf')
```

RMSE, NRMSE & point prediction performance
```{r}
elastic_errors %>%
  ggplot(mapping = aes(x = n, y = RMSE,
                       fill = ticks_included)) +
  geom_bar(stat = 'identity',
           position = position_dodge()) +
  labs(fill = 'Ticks included') +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()) +
  coord_flip() -> p1

elastic_errors %>%
  ggplot(mapping = aes(x = n, y = NRMSE,
                       fill = ticks_included)) +
  geom_bar(stat = 'identity',
           position = position_dodge()) +
  labs(fill = 'Ticks included') +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()) +
  coord_flip() -> p2

elastic_errors %>%
  ggplot(mapping = aes(x = n, y = point_pred_performance,
                       fill = ticks_included)) +
  geom_bar(stat = 'identity',
           position = position_dodge()) +
  labs(fill = 'Ticks included',
       y = 'point prediction performance') +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()) +
  coord_flip() -> p3

p1 + p2 + p3 +
  plot_layout(guides = 'collect') +
  plot_annotation(tag_levels = 'A') & 
  theme(plot.tag = element_text(size = 8))

ggsave('figures/fire_ElasticNet_RMSE.pdf')
```
