---
title: "Fire_03_Create_TrainingData"
author: "Femke Keij S2647168"
date: "2023-08-31"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Clear working directory:
```{r}
rm(list = ls(all = TRUE)) 
```

Packages used:
```{r}
# for importing / working with tidy data
library(tidyverse)

# for arranging ggplots
library(patchwork)

# for computing trends in data
library(trend)
```

Random seed:
```{r}
set.seed(42)
```

Read in the data:
```{r}
# complete data set
fire_full_clean <- read_csv(
  'data/processed/fire_full.csv')

# output data summary
fire_summary_clean <- read_csv(
  'data/processed/fire_summary.csv') %>%
  rename(burn_percentage = mean_burn_percentage)
```

# Mimic real-world data: introduce random noise
ON THE OUTPUT:
Noise is maximum 20% of the mean output value (burn percentage) it is added to (such that smaller values can be as noisy as large ones). How much noise is added is determined by a N(0, 6). 
```{r}
# for the summarised runs
fire_summary_noise <- fire_summary_clean %>%
  # per parameter combination
  group_by(combination_number) %>%
  # compute mean burn percentage over full time series
  mutate(mean_burn_percentage = mean(burn_percentage)) %>%
  # add noise to each value of burn_percentage
  rowwise %>%
  mutate(burn_percentage_noise = burn_percentage +
           mean_burn_percentage * rnorm(n = 1, mean = 0, sd = 6) / 100) %>%
  # add a correction for values < 0, since it's in this case not possible for values to be negative
  mutate(burn_percentage_noise = ifelse(burn_percentage_noise < 0, 0, burn_percentage_noise))

# for the full data
fire_full_noise <- fire_full_clean %>%
  # per parameter combination
  group_by(combination_number, run) %>%
  # compute mean burn percentage over full time series
  mutate(mean_burn_percentage = mean(burn_percentage)) %>%
  # add noise to each value of burn_percentage
  rowwise %>%
  mutate(burn_percentage_noise = burn_percentage +
           mean_burn_percentage * rnorm(n = 1, mean = 0, sd = 6) / 100) %>%
  # add a correction for values < 0, since it's in this case not possible for values to be negative
  mutate(burn_percentage_noise = ifelse(burn_percentage_noise < 0, 0, burn_percentage_noise))
```
Check that the noise was added properly:
```{r}
p1 <- fire_summary_noise %>%
  ggplot(mapping = aes(x = burn_percentage, y = burn_percentage_noise)) +
  geom_point(alpha = 0.5) +
  geom_abline(intercept = 0, slope = 1,
              colour = 'darkgreen', lty = 2) +
  labs(x = 'burn percentage original',
       y = 'burn percentage with noise') +
  theme_minimal() +
  ggtitle('summarised runs')

p2 <- fire_full_noise %>%
  ggplot(mapping = aes(x = burn_percentage, y = burn_percentage_noise)) +
  geom_point(alpha = 0.5) +
  geom_abline(intercept = 0, slope = 1,
              colour = 'darkgreen', lty = 2) +
  labs(x = 'burn percentage original',
       y = 'burn percentage with noise') +
  theme_minimal() +
  ggtitle('full data')

p1 + p2
```

IN TIME:
Time series are shifted by a maximum of 10 time steps. The amount of noise added is sampled randomly from (-10 ; 10).
```{r}
# for the summarised runs
fire_summary_noise <- fire_summary_noise %>%
  group_by(combination_number) %>%
  mutate(ticks_noise = ticks + sample(-10:10, 1))

# for the full data
fire_full_noise <- fire_full_noise %>%
  group_by(combination_number, run) %>%
  mutate(ticks_noise = ticks + sample(-10:10, 1))
```

# Summarise data into time series
Make the following summaries
- End point values with mean, minimum, maximum, sd, and trend of the output;
- Endpoint and middle values, total number of ticks, and mean, minimum, maximum, sd, and trend output;
For the summarised runs and non-summarised runs.

```{r}
TimeSeriesSummarising <- function(data, full = FALSE,
                                  timesteps = TRUE){
  if(full){
    data_return <- data %>%
      # per parameter combination & run number
      group_by(combination_number, run) %>%
      # compute mean, min, max, sd, and trend
      mutate(mean_burn = mean(burn_percentage),
             mean_burn_noise = mean(burn_percentage_noise),
             sd_burn = sd(burn_percentage),
             sd_burn_noise = sd(burn_percentage_noise),
             trend_burn = sens.slope(burn_percentage)$estimates,
             trend_burn_noise =
               sens.slope(burn_percentage_noise)$estimates,
             total_ticks = max(ticks))
  } else {
    data_return <- data %>%
      # per parameter combination
      group_by(combination_number) %>%
      # compute mean, min, max, sd, and trend
      mutate(mean_burn = mean(burn_percentage),
             mean_burn_noise = mean(burn_percentage_noise),
             sd_burn = sd(burn_percentage),
             sd_burn_noise = sd(burn_percentage_noise),
             trend_burn = sens.slope(burn_percentage)$estimates,
             trend_burn_noise =
               sens.slope(burn_percentage_noise)$estimates,
             total_ticks = max(ticks))
  }
  
  if(timesteps){
    data_return <- data_return %>%
      filter(time_label == 'end' | time_label == 'middle')
  } else {
    data_return <- data_return %>%
      filter(time_label == 'end')
  }
  return(data_return)
}
```

Save all data:
```{r}
fire_summary_noise <- fire_summary_noise %>%
  select(- mean_burn_percentage)
fire_full_noise <- fire_full_noise %>%
  select(- mean_burn_percentage)

# summarised runs, timesteps
fire_summary_timesteps <- TimeSeriesSummarising(
  fire_summary_noise, full = FALSE, timesteps = TRUE) %>%
  mutate(sample_size = 198,
         sample_method = 'full',
         summarise_runs = 'summarised',
         datapoints = 'timesteps')

# summarised runs, endpoints only
fire_summary_endpoints <- TimeSeriesSummarising(
  fire_summary_noise, full = FALSE, timesteps = FALSE) %>%
  mutate(sample_size = 198,
         sample_method = 'full',
         summarise_runs = 'summarised',
         datapoints = 'endpoints')

# unsummarised runs, timesteps
fire_full_timesteps <- TimeSeriesSummarising(
  fire_full_noise, full = TRUE, timesteps = TRUE) %>%
  mutate(sample_size = 198,
         sample_method = 'full',
         summarise_runs = 'full',
         datapoints = 'timesteps')

# unsummarised runs, endpoints only
fire_full_endpoints <- TimeSeriesSummarising(
  fire_full_noise, full = TRUE, timesteps = FALSE) %>%
  mutate(sample_size = 198,
         sample_method = 'full',
         summarise_runs = 'full',
         datapoints = 'endpoints')

# merge into 1 file and save as a complete training data set
fire_training <- bind_rows(fire_summary_timesteps,
                           fire_summary_endpoints,
                           fire_full_timesteps,
                           fire_full_endpoints)

fire_training %>%
  write_csv(file = 'data/training/fire_training.csv')
```

# Differentiate
```{r}
fire_training <- read_csv('data/training/fire_training.csv')
```

Retrieve the different training data sets:
- Timesteps, not summarised
```{r}
# retrieve timesteps, non-summarised
data_time_full_clean <- fire_training %>%
  filter(datapoints == 'timesteps',
         summarise_runs == 'full') %>%
  # select relevant variables
  select(time_label, density, directions, burn_percentage, total_ticks,
         mean_burn, sd_burn, trend_burn) %>%
  # switch to wide format so that each timestep is a
  # column (formatted for regression function)
  pivot_wider(names_from = time_label,
              values_from = burn_percentage)

data_time_full_clean %>%
  write_csv('data/training/fire_time_full_clean.csv')

# similar but select noise variables
data_time_full_outputnoise <- fire_training %>%
  filter(datapoints == 'timesteps',
         summarise_runs == 'full') %>%
  select(time_label, density, directions, burn_percentage_noise,
         total_ticks, mean_burn_noise,
         trend_burn_noise, sd_burn_noise) %>%
  # rename so that these can be passed to the regression
  # fitted on the non-noise data
  rename(mean_burn = mean_burn_noise,
         sd_burn = sd_burn_noise,
         trend_burn = trend_burn_noise) %>%
  pivot_wider(names_from = time_label,
              values_from = burn_percentage_noise)

data_time_full_outputnoise %>%
  write_csv('data/training/fire_time_full_outputnoise.csv')
```

Timesteps, summarised
```{r}
data_time_sum_clean <- fire_training %>%
  filter(datapoints == 'timesteps',
         summarise_runs == 'summarised') %>%
  select(time_label, density, directions, burn_percentage, total_ticks,
         mean_burn, sd_burn, trend_burn) %>%
  pivot_wider(names_from = time_label,
              values_from = burn_percentage)

data_time_sum_clean %>%
  write_csv('data/training/fire_time_sum_clean.csv')

data_time_sum_outputnoise <- fire_training %>%
  filter(datapoints == 'timesteps',
         summarise_runs == 'summarised') %>%
  select(time_label, density, directions, burn_percentage_noise,
         total_ticks, mean_burn_noise,
         sd_burn_noise, trend_burn_noise) %>%
  rename(mean_burn = mean_burn_noise,
         sd_burn = sd_burn_noise,
         trend_burn = trend_burn_noise) %>%
  pivot_wider(names_from = time_label,
              values_from = burn_percentage_noise)

data_time_sum_outputnoise %>%
  write_csv('data/training/fire_time_sum_outputnoise.csv')
```

Endpoints, not summarised
```{r}
data_end_full_clean <- fire_training %>%
  filter(datapoints == 'endpoints',
         summarise_runs == 'full') %>%
  select(time_label, density, directions, burn_percentage,
         mean_burn, sd_burn, trend_burn) %>%
  pivot_wider(names_from = time_label,
              values_from = burn_percentage)

data_end_full_clean %>%
  write_csv('data/training/fire_end_full_clean.csv')

data_end_full_outputnoise <- fire_training %>%
  filter(datapoints == 'endpoints',
         summarise_runs == 'full') %>%
  select(time_label, density, directions, burn_percentage_noise,
         mean_burn_noise, sd_burn_noise, trend_burn_noise) %>%
  rename(mean_burn = mean_burn_noise,
         sd_burn = sd_burn_noise,
         trend_burn = trend_burn_noise) %>%
  pivot_wider(names_from = time_label,
              values_from = burn_percentage_noise)

data_end_full_outputnoise %>%
  write_csv('data/training/fire_end_full_outputnoise.csv')
```

Endpoints, summarised
```{r}
data_end_sum_clean <- fire_training %>%
  filter(datapoints == 'endpoints',
         summarise_runs == 'summarised') %>%
  select(time_label, density, directions, burn_percentage,
         mean_burn, sd_burn, trend_burn) %>%
  pivot_wider(names_from = time_label,
              values_from = burn_percentage)

data_end_sum_clean %>%
  write_csv('data/training/fire_end_sum_clean.csv')

data_end_sum_outputnoise <- fire_training %>%
  filter(datapoints == 'endpoints',
         summarise_runs == 'summarised') %>%
  select(time_label, density, directions, burn_percentage_noise,
         mean_burn_noise,
         sd_burn_noise, trend_burn_noise) %>%
  rename(mean_burn = mean_burn_noise,
         sd_burn = sd_burn_noise,
         trend_burn = trend_burn_noise) %>%
  pivot_wider(names_from = time_label,
              values_from = burn_percentage_noise)

data_end_sum_outputnoise %>%
  write_csv('data/training/fire_end_sum_outputnoise.csv')
```