---
title: "WolfSheep_01_Explore_Model"
author: "Femke Keij S2647168"
date: "2023-03-15"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Clear working directory:
```{r}
rm(list = ls(all = TRUE)) 
```

Packages used:
```{r}
# for importing / working with tidy data
library(tidyverse)

# for arranging ggplots
library(patchwork)

# for ggplot aesthetics
library(directlabels)
library(RColorBrewer)

# for computing trends in data
library(trend)
```

Random seed:
```{r}
set.seed(42)
```

Read in the data:
```{r}
# complete data set
ws_output <- read_csv(
  'data/raw/wolfsheep_output.csv')

# training data
#ws_train <- read_csv(
#  'data/processed/wolfsheep_train.csv')

# test data
#ws_test <- read_csv(
#  'data/processed/wolfsheep_test.csv')

# output data summary
ws_summary <- read_csv(
  'data/processed/wolfsheep_summary.csv')

# parameter combinations
#all_comb <- read_csv('data/processed/wolfsheep_parameterspace.csv')
```

# Inspecting parameter space: looking for equifinality
Initial sheep and wolves plus the resulting final average numbers of sheep and wolves
```{r}
ws_summary_lasttick <- ws_summary %>%
  group_by(combination_number) %>%
  mutate(max_ticks = max(ticks)) %>%
  filter(ticks == max_ticks) %>%
  ungroup()

p1 <- ws_summary_lasttick %>%
  ggplot(mapping = aes(x = initial_number_sheep, y = initial_number_wolves)) +
  geom_point(mapping = aes(colour = mean_sheep),
             size = 5) +
  theme_minimal()+
  theme(panel.grid.minor = element_blank()) +
  labs(x = 'initial number of sheep',
       y = 'initial number of wolves',
       colour = 'mean number of \n sheep at last tick') +
  scale_colour_gradient(low = 'navy', high = 'red')

p2 <- ws_summary_lasttick %>%
  ggplot(mapping = aes(x = initial_number_sheep, y = initial_number_wolves)) +
  geom_point(mapping = aes(colour = mean_wolves),
             size = 5) +
  theme_minimal()+
  theme(panel.grid.minor = element_blank()) +
  labs(x = 'initial number of sheep',
       y = 'initial number of wolves',
       colour = 'mean number of \n wolves at last tick') +
  scale_colour_gradient(low = 'navy', high = 'red')

p1 + p2
```

Visualise all time series:
```{r}
p_sheep <- ws_summary %>%
  ggplot(mapping = aes(x = ticks, y = mean_sheep,
                       group = combination_number)) +
  geom_line(colour = 'bisque2') +
  theme_minimal() +
  labs(y = 'sheep')

p_wolves <- ws_summary %>%
  ggplot(mapping = aes(x = ticks, y = mean_wolves,
                       group = combination_number)) +
  geom_line(colour = 'darkgrey') +
  theme_minimal() +
  labs(y = 'wolves')

p_grass <- ws_summary %>%
  ggplot(mapping = aes(x = ticks, y = mean_grass,
                       group = combination_number)) +
  geom_line(colour = 'darkgreen') +
  theme_minimal() +
  labs(y = 'grass')

p_sheep + p_wolves + p_grass
```

Now do the same, but color each based on the values of the 7 parameters:
```{r}
ws_summary_cat <- ws_summary %>%
  mutate(sheep_cat = cut(initial_number_sheep,
                         breaks = c(-Inf, 50, 100, 150, 200, Inf),
                         labels = c('0 - 50', '51 - 100', '101 - 150',
                                    '151 - 200', '201 - 250')),
         wolves_cat = cut(initial_number_wolves,
                          breaks = c(-Inf, 50, 100, 150, 200, Inf),
                          labels = c('0 - 50', '51 - 100', '101 - 150',
                                    '151 - 200', '201 - 250')),
         sheep_food_cat = cut(sheep_gain_from_food,
                         breaks = c(-Inf, 25, Inf),
                         labels = c('0 - 25', '26 - 50')),
         wolves_food_cat = cut(wolves_gain_from_food,
                         breaks = c(-Inf, 25, 50, 75, Inf),
                         labels = c('0 - 25', '26 - 50', '51 - 75', '76 - 100')),
         sheep_rep_cat = cut(sheep_reproduce,
                         breaks = c(-Inf, 10, Inf),
                         labels = c('0 - 10', '11 - 20')),
         wolves_rep_cat = cut(wolves_reproduce,
                         breaks = c(-Inf, 10, Inf),
                         labels = c('1 - 10', '11 - 20')),
         grass_cat = cut(grass_regrowth,
                         breaks = c(-Inf, 25, 50, 75, Inf),
                         labels = c('0 - 25', '26 - 50', '51 - 75', '76 - 100')))
```

Function for plots
```{r}
PlotParameterSpace <- function(parameter, label){
  p_sheep <- ws_summary_cat %>%
    ggplot(mapping = aes(x = ticks, y = mean_sheep,
                       group = combination_number, colour = get(parameter))) +
    geom_line() +
    theme_minimal() +
    labs(y = 'sheep',
         colour = label)

  p_wolves <- ws_summary_cat %>%
    ggplot(mapping = aes(x = ticks, y = mean_wolves,
                       group = combination_number, colour = get(parameter))) +
    geom_line() +
    theme_minimal() +
    labs(y = 'wolves',
         colour = label)

  p_grass <- ws_summary_cat %>%
    ggplot(mapping = aes(x = ticks, y = mean_grass,
                       group = combination_number, colour = get(parameter))) +
    geom_line() +
    theme_minimal() +
    labs(y = 'grass',
         colour = label)
  
  plot <- p_sheep + p_wolves + p_grass + plot_layout(guides = 'collect')
  
  return(plot)
}
```

1) initial_number_sheep
```{r}
PlotParameterSpace(parameter = 'sheep_cat',
                   label = 'initial number of sheep')
```

2) initial_number_wolves
```{r}
PlotParameterSpace(parameter = 'wolves_cat',
                   label = 'initial numbr of wolves')
```

3) sheep_gain_from_food
```{r}
PlotParameterSpace(parameter = 'sheep_food_cat',
                   label = 'sheep gain from food')
```

4) wolves_gain_from_food
```{r}
PlotParameterSpace(parameter = 'wolves_food_cat',
                   label = 'wolves gain from food')
```

5) sheep_reproduce
```{r}
PlotParameterSpace(parameter = 'sheep_rep_cat',
                   label = 'probability sheep reproduce')
```

6) wolves_reproduce
```{r}
PlotParameterSpace(parameter = 'wolves_rep_cat',
                   label = 'probability wolves reproduce')
```

7) grass_regrowth_time
```{r}
PlotParameterSpace(parameter = 'grass_cat',
                   label = 'grass regrowth time')
```

It seems very difficult to say anything about equifinality here, since there are so many different parameters to look at at once.

# Inspecting parameter space: Identifying unidentifiable parameters
Correlation between parameters and outputs:
```{r}
PlotCorrelations <- function(parameter){
  p_sheep <- ws_summary %>%
    filter(ticks %in% c(50, 100, 250, 500)) %>%
    ggplot(mapping = aes(x = get(parameter), y = mean_sheep)) +
    geom_point() +
    geom_smooth(method = 'lm', se = FALSE) +
    theme_minimal() +
    labs(x = parameter, y = 'sheep') +
    facet_grid(~ as.factor(ticks))
  
  p_wolves <- ws_summary %>%
    filter(ticks %in% c(50, 100, 250, 500)) %>%
    ggplot(mapping = aes(x = get(parameter), y = mean_wolves)) +
    geom_point() +
    geom_smooth(method = 'lm', se = FALSE) +
    theme_minimal() +
    labs(x = parameter, y = 'wolves') +
    facet_grid(~ as.factor(ticks))
  
  p_grass <- ws_summary %>%
    filter(ticks %in% c(50, 100, 250, 500)) %>%
    ggplot(mapping = aes(x = get(parameter), y = mean_grass)) +
    geom_point() +
    geom_smooth(method = 'lm', se = FALSE) +
    theme_minimal() +
    labs(x = parameter, y = 'grass') +
    facet_grid(~ as.factor(ticks))
  
  plot <- p_sheep / p_wolves / p_grass + plot_layout(guides = 'collect')
  
  return(plot)
}
```

1) initial_number_sheep
```{r}
PlotCorrelations('initial_number_sheep')
```

2) initial_number_wolves
```{r}
PlotCorrelations('initial_number_wolves')
```

3) sheep_gain_from_food
```{r}
PlotCorrelations('sheep_gain_from_food')
```

4) wolves_gain_from_food
```{r}
PlotCorrelations('wolves_gain_from_food')
```

5) sheep_reproduce
```{r}
PlotCorrelations('sheep_reproduce')
```

6) wolves_reproduce
```{r}
PlotCorrelations('wolves_reproduce')
```

7) grass_regrowth_time
```{r}
PlotCorrelations('grass_regrowth')
```

# Mimic real-world data: introduce random noise
Noise is maximum 20% of the mean output value it is added to (such that smaller values can be as noisy as large ones). How much noise is added is determined by a N(0, 6). 
```{r}
ws_summary <- ws_summary %>%
  rename(sheep = mean_sheep,
         wolves = mean_wolves,
         grass = mean_grass) %>%
  # per parameter combination
  group_by(combination_number) %>%
  # compute mean sheep, grass, wolves over full time series
  mutate(mean_sheep = mean(sheep),
         mean_wolves = mean(wolves),
         mean_grass = mean(grass)) %>%
  # add noise to each value of sheep, grass, and wolves
  rowwise %>%
  mutate(sheep_noise = sheep + mean_sheep * rnorm(n = 1, mean = 0, sd = 6) / 100,
         wolves_noise = wolves + mean_wolves * rnorm(n = 1, mean = 0, sd = 6) / 100,
         grass_noise = grass + mean_grass * rnorm(n = 1, mean = 0, sd = 6) / 100) %>%
  # add a correction for values < 0, since it's in this case not possible for values
  # to be negative
  mutate(sheep_noise = ifelse(sheep_noise < 0, 0, sheep_noise),
         wolves_noise = ifelse(wolves_noise < 0, 0, wolves_noise),
         grass_noise = ifelse(grass_noise < 0, 0, grass_noise))
```
Check that the noise was added properly:
```{r}
# sheep
ws_summary %>%
  ggplot(mapping = aes(x = sheep, y = sheep_noise)) +
  geom_point(alpha = 0.5) +
  geom_abline(intercept = 0, slope = 1,
              colour = 'darkgreen', lty = 2) +
  labs(x = 'sheep original', y = 'sheep noise') +
  theme_minimal()

# wolves
ws_summary %>%
  ggplot(mapping = aes(x = wolves, y = wolves_noise)) +
  geom_point(alpha = 0.5) +
  geom_abline(intercept = 0, slope = 1,
              colour = 'darkgreen', lty = 2) +
  labs(x = 'wolves original', y = 'wolves noise') +
  theme_minimal()

# grass
ws_summary %>%
  ggplot(mapping = aes(x = grass, y = grass_noise)) +
  geom_point(alpha = 0.5) +
  geom_abline(intercept = 0, slope = 1,
              colour = 'darkgreen', lty = 2) +
  labs(x = 'grass original', y = 'grass noise') +
  theme_minimal()
```

# Summarise time series:
Carella 2021 does the following for this model (with 200 time steps):
- value every 50 time steps
- maximum
- minimum
- average
- standard deviation
- trend
I'm going to add an option with
- value every 20 time steps
- maximum
- minimum
- average
- standard deviation
- trend
And an option with
- value every 10 time steps
- maximum
- minimum
- average
- standard deviation
- trend
And also an option where we look at the entire time series
+ looking only at the final time point.

```{r}
ws_summary_timesteps <- ws_summary %>%
  # per parameter combination
  group_by(combination_number) %>%
  # compute min, max, sd, and trend for sheep, grass, wolves over full time series
  # do this for both the clean and the noisy data
  mutate(min_sheep = min(sheep),
         max_sheep = max(sheep),
         sd_sheep = sd(sheep),
         trend_sheep = sens.slope(sheep)$estimates,
         mean_sheep_noise = mean(sheep_noise),
         min_sheep_noise = min(sheep_noise),
         max_sheep_noise = max(sheep_noise),
         sd_sheep_noise = sd(sheep_noise),
         trend_sheep_noise = sens.slope(sheep_noise)$estimates,
         min_wolves = min(wolves),
         max_wolves = max(wolves),
         sd_wolves = sd(wolves),
         trend_wolves = sens.slope(wolves)$estimates,
         mean_wolves_noise = mean(wolves_noise),
         min_wolves_noise = min(wolves_noise),
         max_wolves_noise = max(wolves_noise),
         sd_wolves_noise = sd(wolves_noise),
         trend_wolves_noise = sens.slope(wolves_noise)$estimates,
         min_grass = min(grass),
         max_grass = max(grass),
         sd_grass = sd(grass),
         trend_grass = sens.slope(grass)$estimates,
         mean_grass_noise = mean(grass_noise),
         min_grass_noise = min(grass_noise),
         max_grass_noise = max(grass_noise),
         sd_grass_noise = sd(grass_noise),
         trend_grass_noise = sens.slope(grass_noise)$estimates) %>%
  ungroup() %>%
  # filter out the time steps of interest
  filter(ticks %in% seq(from = 10, to = 500, by = 10))
```
Reshape data
```{r}
# extend all data to 500 ticks

# new data frame to store extended data
ws_summary_timesteps_new <- ws_summary_timesteps[1,]

# per parameter combination
for(i in unique(ws_summary_timesteps$combination_number)){
  # extract data for parameter combination
  work_with <- ws_summary_timesteps %>%
    filter(combination_number == i)
  
  # if data contains less than 500 ticks
  if(max(work_with$ticks) < 500){
    # register last tick included
    max_ticks <- max(work_with$ticks)
    
    # data to extend to 500 ticks
    add_rows <- work_with %>%
      filter(ticks == max_ticks)
    
    # ticks to add
    add_ticks <- seq(from = max_ticks + 10, to = 500, by = 10)
    
    # to store new data
    new_df <- add_rows
    
    # create data frame with last tick data, extending
    # for all required ticks
    for(i in 1:length(add_ticks)){
      new_df <- bind_rows(new_df, add_rows)
    }
    
    # set ticks to run until 500
    new_df$ticks <- c(max_ticks, add_ticks)
    # add original data to extended data
    new_df <- bind_rows(work_with, new_df)
    # to send back to complete dataframe
    work_with <- new_df
  }
  # fill out new data frame with extended data
  # or original data where it didn't need to be extended
  ws_summary_timesteps_new <- bind_rows(ws_summary_timesteps_new, work_with)
}
# remove set-up data row
ws_summary_timesteps_new <- ws_summary_timesteps_new[-1, ] %>%
  distinct()

# pivot wider so that each time step is a data column
# each parameter combination will have 1 row of data
ws_summary_timesteps <- ws_summary_timesteps_new %>%
  filter(ticks %in% seq(from = 10, to = 500, by = 10)) %>%
  pivot_wider(names_from = ticks,
              values_from = c(sheep, wolves, grass,
                              sheep_noise, wolves_noise, grass_noise))
```

Save all data frames:
```{r}
# complete data (for time series analysis)
ws_summary %>%
  write_csv(file = 'data/processed/wolfsheep_timeseries_training.csv')
# time steps
ws_summary_timesteps %>%
  write_csv(file = 'data/processed/wolfsheep_timesteps_training.csv')
```