---
title: "WolfSheep_00_Run_Model"
author: "Femke Keij S2647168"
date: "2023-03-13"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import Python modules:
```{python}
import pandas as pd
import matplotlib.pyplot as plt
import seaborn as sns
import math
import statistics
import itertools
sns.set_style('white')
sns.set_context('talk')

import pyNetLogo # to run NetLogo from RStudio
```

Import R packages:
```{r}
# for importing / working with tidy data
library(tidyverse)
# for arranging plots
library(patchwork)
# for cumulative computations
library(zoo)
# for rolling averages
library(slider)
# for latin hypercube sampling
library(lhs)
```

Set seed:
```{r}
set.seed(42)
```

Link to and start the wolf-sheep-grass model:
```{python}
netlogo = pyNetLogo.NetLogoLink(gui = True) # start netlogo

netlogo.load_model(r"C:\Users\Femke Keij\OneDrive\Thesis\R projects\LU thesis local\models\WolfSheep_myversion.nlogo")
# open model

# set to correct model version
netlogo.command('set model-version "sheep-wolves-grass"')
```

# Decide on model outputs
I'll use the number of sheep, number of wolves, and amount of grass as outputs. I'm recording them at each tick for 500 ticks.

# Determine number of runs
1) coefficient of variation
2) mean convergence
3) windowed variance

Optionally, we can set a separate convergence boundary $\epsilon$ for each output. This is only necessary when the outputs are on different scales, which is not the case here. I choose $\epsilon_{c_\upsilon} = 0.1$, $\epsilon_\mu = 1$, and $\epsilon_{\omega^2} = 0.2$.

We can compute all these statistics using just 1 data set. To do so, we run each parameter combination 50 times and save the outputs at time step 100 (this is an arbitrary choice that should potentially be remedied later to include the time steps etc. that I'm actually going to use).

Instead of combining a low, mid, and high value for each parameter, I use only the 25% and 75% quartile of each range. This brings the number of parameter combinations down from 2187 to 128.
```{python}
initial_number_sheep = [62, 188]
initial_number_wolves = [62, 188]
sheep_gain_from_food = [12, 38]
wolf_gain_from_food = [25, 72]
sheep_reproduce = [5, 15]
wolf_reproduce = [5, 15]
grass_regrowth = [25, 72]

dictionary = {'initial_number_sheep': [62, 188], 'initial_number_wolves': [62, 188], 'sheep_gain_from_food': [12, 38], 'wolf_gain_from_food': [25, 72], 'sheep_reproduce': [5, 15], 'wolf_reproduce': [5, 15], 'grass_regrowth': [25, 72]}

def expand_grid(data_dict):
    rows = itertools.product(*data_dict.values())
    return pd.DataFrame.from_records(rows, columns = data_dict.keys())

param_comb = expand_grid(dictionary)
```

Obtain all necessary data:
```{python}
# data frame to store results
out_df = pd.DataFrame(columns = ['initial_number_sheep', 'initial_number_wolves', 'sheep_gain_from_food', 'wolf_gain_from_food', 'sheep_reproduce', 'wolf_reproduce', 'grass_regrowth', 'run', 'wolves', 'sheep', 'grass', 'tick'])
# add empty row to start
out_df = out_df.append({'initial_number_sheep': 0, 'initial_number_wolves': 0, 'sheep_gain_from_food': 0, 'wolf_gain_from_food': 0, 'sheep_reproduce': 0, 'wolf_reproduce': 0, 'grass_regrowth': 0, 'run': 0, 'sheep': -100, 'wolves': -100, 'grass': -100, 'tick': - 100}, ignore_index = True)

# set counter
count = 0
# set random seed
netlogo.command('clean')

for i in range(0,len(param_comb)):
  # set parameters
  initial_number_sheep = str(param_comb['initial_number_sheep'][i])
  initial_number_wolves = str(param_comb['initial_number_wolves'][i])
  sheep_gain_from_food = str(param_comb['sheep_gain_from_food'][i])
  wolf_gain_from_food = str(param_comb['wolf_gain_from_food'][i])
  sheep_reproduce = str(param_comb['sheep_reproduce'][i])
  wolf_reproduce = str(param_comb['wolf_reproduce'][i])
  grass_regrowth = str(param_comb['grass_regrowth'][i])
  netlogo.command(''.join(['set initial-number-sheep ', initial_number_sheep]))
  netlogo.command(''.join(['set initial-number-wolves ', initial_number_wolves]))
  netlogo.command(''.join(['set  sheep-gain-from-food ',  sheep_gain_from_food]))
  netlogo.command(''.join(['set wolf-gain-from-food ',  wolf_gain_from_food]))
  netlogo.command(''.join(['set sheep-reproduce ', sheep_reproduce]))
  netlogo.command(''.join(['set wolf-reproduce ', wolf_reproduce]))
  netlogo.command(''.join(['set grass-regrowth-time ', grass_regrowth]))
  
  for x in range(0, 50):
    # setup
    netlogo.command('setup')
    # execute for 50 ticks, 3 times
    for j in [50, 100, 150]:
      netlogo.command('repeat 50 [go]')
      # record sheep, wolves, grass
      sheep = netlogo.report('count sheep')
      wolves = netlogo.report('count wolves')
      grass = netlogo.report ('count grass')
      # store
      out_df = out_df.append({'initial_number_sheep': initial_number_sheep, 'initial_number_wolves': initial_number_wolves, 'sheep_gain_from_food': sheep_gain_from_food, 'wolf_gain_from_food': wolf_gain_from_food, 'sheep_reproduce': sheep_reproduce, 'wolf_reproduce': wolf_reproduce, 'grass_regrowth': grass_regrowth, 'run': x + 1, 'sheep': sheep, 'wolves': wolves, 'grass': grass, 'tick': j}, ignore_index = True)

out_df.to_csv('data/raw/wolfsheep_numruns.csv')
```

Compute statistics:
```{r}
numruns <- read.csv('data/raw/wolfsheep_numruns.csv')

# remove first row
numruns <- numruns %>%
  slice(-1)

# add a unique number for each parameter combination
numruns <- numruns %>%
  group_by(initial_number_sheep, initial_number_wolves, sheep_gain_from_food,
           wolf_gain_from_food, sheep_reproduce, wolf_reproduce, grass_regrowth) %>%
  mutate(combination_number = cur_group_id())

# pivot to wide format
numruns_wide <- numruns %>%
  select(- X) %>%
  pivot_wider(names_from = tick,
              values_from = c(sheep, wolves, grass))
```

1) Mean convergence
With $\epsilon_\mu = 1$
```{r}
numruns <- numruns %>%
  group_by(combination_number, tick) %>%
  # compute cumulative means
  # and differences between means
  mutate(mean_sheep = cummean(sheep),
         mean_wolves = cummean(wolves),
         mean_grass = cummean(grass),
         mean_diff_sheep = mean_sheep - 
           lag(mean_sheep, default = mean_sheep[1]),
         mean_diff_wolves = mean_wolves -
           lag(mean_wolves, default = mean_wolves[1]),
         mean_diff_grass = mean_grass -
           lag(mean_grass, default = mean_grass[1]))

# compute at which point the mean difference settles below 1
mean_conv_sheep <- numruns %>%
  group_by(combination_number, tick) %>%
  filter(abs(mean_diff_sheep) < 1,
         run != 1) %>%
  slice(which.min(run))

mean_conv_wolves <- numruns %>%
  group_by(combination_number, tick) %>%
  filter(abs(mean_diff_wolves) < 1,
         run != 1) %>%
  slice(which.min(run))

mean_conv_grass <- numruns %>%
  group_by(combination_number, tick) %>%
  filter(abs(mean_diff_grass) < 1,
         run != 1) %>%
  slice(which.min(run))

# plot for sheep
p1_sheep <- numruns %>%
  ggplot(mapping = aes(x = run, y = mean_diff_sheep,
                       group = combination_number)) +
  geom_line(alpha = 0.5) +
  geom_hline(yintercept = 1, colour = 'darkgreen') +
  geom_hline(yintercept = - 1, colour = 'darkgreen') +
  labs(y = 'difference mean number',
       title = 'sheep')

p2_sheep <- mean_conv_sheep %>%
  ggplot(mapping = aes(x = run)) +
  geom_histogram(binwidth = 1) +
  geom_vline(xintercept = mean(mean_conv_sheep$run),
             colour = 'darkgreen', lty = 2) +
  geom_vline(xintercept = median(mean_conv_sheep$run),
             colour = 'darkblue', lty = 2) +
  annotate(geom = 'text', label = 'mean',
           x = mean(mean_conv_sheep$run) + 1, y = 50,
           colour = 'darkgreen', angle = 90) +
  annotate(geom = 'text', label = 'median',
           x = median(mean_conv_sheep$run) + 1, y = 50,
           colour = 'darkblue', angle = 90) +
  labs(y = 'count')

# plot for wolves
p1_wolves <- numruns %>%
  ggplot(mapping = aes(x = run, y = mean_diff_wolves,
                       group = combination_number)) +
  geom_line(alpha = 0.5) +
  geom_hline(yintercept = 1, colour = 'darkgreen') +
  geom_hline(yintercept = - 1, colour = 'darkgreen') +
  labs(x = 'runs added',
       y = 'difference mean number',
       title = 'wolves')

p2_wolves <- mean_conv_wolves %>%
  ggplot(mapping = aes(x = run)) +
  geom_histogram(binwidth = 1) +
  geom_vline(xintercept = mean(mean_conv_wolves$run),
             colour = 'darkgreen', lty = 2) +
  geom_vline(xintercept = median(mean_conv_wolves$run),
             colour = 'darkblue', lty = 2) +
  annotate(geom = 'text', label = 'mean',
           x = mean(mean_conv_wolves$run) + 1, y = 40,
           colour = 'darkgreen', angle = 90) +
  annotate(geom = 'text', label = 'median',
           x = median(mean_conv_wolves$run) - 1, y = 40,
           colour = 'darkblue', angle = 90) +
  labs(y = 'count',
       x = 'number of runs until mean convergence')

# plot for grass
p1_grass <- numruns %>%
  ggplot(mapping = aes(x = run, y = mean_diff_grass,
                       group = combination_number)) +
  geom_line(alpha = 0.5) +
  geom_hline(yintercept = 1, colour = 'darkgreen') +
  geom_hline(yintercept = - 1, colour = 'darkgreen') +
  labs(y = 'difference mean number',
       title = 'grass')

p2_grass <- mean_conv_grass %>%
  ggplot(mapping = aes(x = run)) +
  geom_histogram(binwidth = 1) +
  geom_vline(xintercept = mean(mean_conv_grass$run),
             colour = 'darkgreen', lty = 2) +
  geom_vline(xintercept = median(mean_conv_grass$run),
             colour = 'darkblue', lty = 2) +
  annotate(geom = 'text', label = 'mean',
           x = mean(mean_conv_grass$run) + 1, y = 40,
           colour = 'darkgreen', angle = 90) +
  annotate(geom = 'text', label = 'median',
           x = median(mean_conv_grass$run) - 1, y = 40,
           colour = 'darkblue', angle = 90) +
  labs(y = 'count')

# plot all together
mean_conv_plot <- (p1_sheep + p1_wolves + p1_grass) /
  (p2_sheep + p2_wolves + p2_grass) +
  plot_layout(guides = 'collect') +
  plot_annotation(tag_levels = '1') &
  theme_minimal() +
  theme(plot.tag = element_text(size = 8),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(fill = NA, colour = 'lightgrey'))

mean_conv_plot[[1]][[1]] <- mean_conv_plot[[1]][[1]] +
  theme(axis.title.x = element_blank())
mean_conv_plot[[1]][[2]] <- mean_conv_plot[[1]][[2]] +
  theme(axis.title.y = element_blank())
mean_conv_plot[[1]][[3]] <- mean_conv_plot[[1]][[3]] +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())

mean_conv_plot[[2]][[1]] <- mean_conv_plot[[2]][[1]] +
  theme(axis.title.x = element_blank())
mean_conv_plot[[2]][[2]] <- mean_conv_plot[[2]][[2]] +
  theme(axis.title.y = element_blank())
mean_conv_plot[[2]][[3]] <- mean_conv_plot[[2]][[3]] +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())
```

2) Coefficient of variation
With $\epsilon_{c_\upsilon} = 0.1$.
```{r}
numruns <- numruns %>%
  group_by(combination_number) %>%
  # compute cumulative standard deviations
  # compute coefficient of variation
  mutate(sd_sheep = rollapplyr(sheep, 1:n(), sd),
         sd_wolves = rollapplyr(wolves, 1:n(), sd),
         sd_grass = rollapplyr(grass, 1:n(), sd),
         cv_sheep = sd_sheep / mean_sheep,
         cv_wolves = sd_wolves / mean_wolves,
         cv_grass = sd_grass / mean_grass) %>%
  # replace all Inf with NA (resulting from mean = 0)
  mutate(cv_sheep = ifelse(abs(cv_sheep) == Inf, NA, cv_sheep),
         cv_wolves = ifelse(abs(cv_wolves) == Inf, NA, cv_wolves),
         cv_grass = ifelse(abs(cv_grass) == Inf, NA, cv_grass)) %>%
  # replace all NAs (resulting from sd = 0 or mean = 0) with 0
  replace(is.na(.), 0) %>%
  # compute differences in cvs
  mutate(cv_diff_sheep = cv_sheep - lag(cv_sheep,
                                        default = cv_sheep[1]),
         cv_diff_wolves = cv_wolves - lag(cv_wolves,
                                          default = cv_wolves[1]),
         cv_diff_grass = cv_grass - lag(cv_grass,
                                        default = cv_grass[1]))
  
# compute at which point the mean difference settles below 1
cv_conv_sheep <- numruns %>%
  group_by(combination_number) %>%
  filter(abs(cv_diff_sheep) < 0.1,
         run != 1) %>%
  slice(which.min(run))

cv_conv_wolves <- numruns %>%
  group_by(combination_number) %>%
  filter(abs(cv_diff_wolves) < 0.1,
         run != 1) %>%
  slice(which.min(run))

cv_conv_grass <- numruns %>%
  group_by(combination_number) %>%
  filter(abs(cv_diff_grass) < 0.1,
         run != 1) %>%
  slice(which.min(run))

# plot for sheep
p1_sheep <- numruns %>%
  ggplot(mapping = aes(x = run, y = cv_diff_sheep,
                       group = combination_number)) +
  geom_line(alpha = 0.5) +
  geom_hline(yintercept = 0.1, colour = 'darkgreen') +
  geom_hline(yintercept = - 0.1, colour = 'darkgreen') +
  labs(x = 'added runs',
       y = 'cv',
       title = 'sheep') +
  facet_grid(~ tick)

p2_sheep <- cv_conv_sheep %>%
  ggplot(mapping = aes(x = run)) +
  geom_histogram(binwidth = 1) +
  geom_vline(xintercept = mean(cv_conv_sheep$run),
             colour = 'darkgreen', lty = 2) +
  geom_vline(xintercept = median(cv_conv_sheep$run),
             colour = 'darkblue', lty = 2) +
  annotate(geom = 'text', label = 'mean',
           x = mean(cv_conv_sheep$run) + 1, y = 20,
           colour = 'darkgreen', angle = 90) +
  annotate(geom = 'text', label = 'median',
           x = median(cv_conv_sheep$run) - 1.5, y = 20,
           colour = 'darkblue', angle = 90) +
  labs(y = 'count',
       x = 'number of runs until cv convergence')

# plot for wolves
p1_wolves <- numruns %>%
  ggplot(mapping = aes(x = run, y = cv_diff_wolves,
                       group = combination_number)) +
  geom_line(alpha = 0.5) +
  geom_hline(yintercept = 0.1, colour = 'darkgreen') +
  geom_hline(yintercept = - 0.1, colour = 'darkgreen') +
  labs(x = 'added runs',
       y = 'difference cv',
       title = 'wolves') +
  facet_grid(~ tick)

p2_wolves <- cv_conv_wolves %>%
  ggplot(mapping = aes(x = run)) +
  geom_histogram(binwidth = 1) +
  geom_vline(xintercept = mean(cv_conv_wolves$run),
             colour = 'darkgreen', lty = 2) +
  geom_vline(xintercept = median(cv_conv_wolves$run),
             colour = 'darkblue', lty = 2) +
  annotate(geom = 'text', label = 'mean',
           x = mean(cv_conv_wolves$run) + 1, y = 20,
           colour = 'darkgreen', angle = 90) +
  annotate(geom = 'text', label = 'median',
           x = median(cv_conv_wolves$run) - 1.5, y = 20,
           colour = 'darkblue', angle = 90) +
  labs(y = 'count',
       x = 'number of runs until cv convergence')

# plot for grass
p1_grass <- numruns %>%
  ggplot(mapping = aes(x = run, y = cv_diff_grass,
                       group = combination_number)) +
  geom_line(alpha = 0.5) +
  geom_hline(yintercept = 0.1, colour = 'darkgreen') +
  geom_hline(yintercept = - 0.1, colour = 'darkgreen') +
  labs(x = 'added runs',
       y = 'difference cv',
       title = 'grass') +
  facet_grid(~ tick)

p2_grass <- cv_conv_grass %>%
  ggplot(mapping = aes(x = run)) +
  geom_histogram(binwidth = 1) +
  geom_vline(xintercept = mean(cv_conv_grass$run),
             colour = 'darkgreen', lty = 2) +
  geom_vline(xintercept = median(cv_conv_grass$run),
             colour = 'darkblue', lty = 2) +
  annotate(geom = 'text', label = 'mean',
           x = mean(cv_conv_grass$run) + 1, y = 20,
           colour = 'darkgreen', angle = 90) +
  annotate(geom = 'text', label = 'median',
           x = median(cv_conv_grass$run) - 1, y = 20,
           colour = 'darkblue', angle = 90) +
  labs(y = 'count',
       x = 'number of runs until cv convergence')

# plot all together
cv_conv_plot <- (p1_sheep + p1_wolves + p1_grass) /
  (p2_sheep + p2_wolves + p2_grass) +
  plot_layout(guides = 'collect') +
  plot_annotation(tag_levels = '1') &
  theme_minimal() +
  theme(plot.tag = element_text(size = 8),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(fill = NA, colour = 'lightgrey'))

cv_conv_plot[[1]][[1]] <- cv_conv_plot[[1]][[1]] +
  theme(axis.title.x = element_blank())
cv_conv_plot[[1]][[2]] <- cv_conv_plot[[1]][[2]] +
  theme(axis.title.y = element_blank())
cv_conv_plot[[1]][[3]] <- cv_conv_plot[[1]][[3]] +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())

cv_conv_plot[[2]][[1]] <- cv_conv_plot[[2]][[1]] +
  theme(axis.title.x = element_blank())
cv_conv_plot[[2]][[2]] <- cv_conv_plot[[2]][[2]] +
  theme(axis.title.y = element_blank())
cv_conv_plot[[2]][[3]] <- cv_conv_plot[[2]][[3]] +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())
```

3) Windowed variance / relative outer variance
Width of window: 10 runs
With $\epsilon_\omega^2 = 0.2$
```{r}
numruns <- numruns %>% 
  group_by(combination_number) %>% 
  mutate(windowed_var_sheep = slider::slide_dbl(sheep, var, .before = 5, .after = 5),
         windowed_var_wolves = slider::slide_dbl(wolves, var, .before = 5, .after = 5),
         windowed_var_grass = slider::slide_dbl(grass, var, .before = 5, .after = 5),
         relative_outer_var_sheep = windowed_var_sheep / max(windowed_var_sheep),
         relative_outer_var_wolves = windowed_var_wolves / max(windowed_var_wolves),
         relative_outer_var_grass = windowed_var_grass / max(windowed_var_grass)) %>%
  # there's a few instances where the relative outer variance is NaN because the population
  # is 0, therefore, we set those to 0 instead of NaN
  mutate_at(vars(relative_outer_var_sheep,
                 relative_outer_var_grass,
                 relative_outer_var_wolves), ~ replace(., is.nan(.), 0))

# compute at which point the relative outer variance settles below 0.2
omega_conv_sheep <- numruns %>%
  group_by(combination_number) %>%
  filter(abs(relative_outer_var_sheep) < 0.2,
         run != 1) %>%
  slice(which.min(run))

omega_conv_wolves <- numruns %>%
  group_by(combination_number) %>%
  filter(abs(relative_outer_var_wolves) < 0.2,
         run != 1) %>%
  slice(which.min(run))

omega_conv_grass <- numruns %>%
  group_by(combination_number) %>%
  filter(abs(relative_outer_var_grass) < 0.2,
         run != 1) %>%
  slice(which.min(run)) 

# plot for sheep
p1_sheep <- numruns %>%
  ggplot(mapping = aes(x = run, y = relative_outer_var_sheep,
                       group = combination_number)) +
  geom_line(alpha = 0.2) +
  geom_hline(yintercept = 0.2, colour = 'darkgreen') +
  labs(y = 'relative outer variance sheep',
       title = 'sheep') +
  facet_grid(~ tick) +
  theme(axis.title.x = NULL)

p2_sheep <- omega_conv_sheep %>%
  ggplot(mapping = aes(x = run)) +
  geom_histogram(binwidth = 1) +
  geom_vline(xintercept = mean(omega_conv_sheep$run),
             colour = 'darkgreen', lty = 2) +
  geom_vline(xintercept = median(omega_conv_sheep$run),
             colour = 'darkblue', lty = 2) +
  annotate(geom = 'text', label = 'mean',
           x = mean(omega_conv_sheep$run) + 1, y = 20,
           colour = 'darkgreen', angle = 90) +
  annotate(geom = 'text', label = 'median',
           x = median(omega_conv_sheep$run) - 1.5, y = 20,
           colour = 'darkblue', angle = 90) +
  labs(y = 'count') +
  theme(axis.title.x = NULL)

# plot for wolves
p1_wolves <- numruns %>%
  ggplot(mapping = aes(x = run, y = relative_outer_var_wolves,
                       group = combination_number)) +
  geom_line(alpha = 0.2) +
  geom_hline(yintercept = 0.2, colour = 'darkgreen') +
  labs(y = 'relative outer variance wolves',
       title = 'wolves') +
  facet_grid(~ tick)

p2_wolves <- omega_conv_wolves %>%
  ggplot(mapping = aes(x = run)) +
  geom_histogram(binwidth = 1) +
  geom_vline(xintercept = mean(omega_conv_wolves$run),
             colour = 'darkgreen', lty = 2) +
  geom_vline(xintercept = median(omega_conv_wolves$run),
             colour = 'darkblue', lty = 2) +
  annotate(geom = 'text', label = 'mean',
           x = mean(omega_conv_wolves$run) + 1, y = 10,
           colour = 'darkgreen', angle = 90) +
  annotate(geom = 'text', label = 'median',
           x = median(omega_conv_wolves$run) + 1, y = 10,
           colour = 'darkblue', angle = 90) +
  labs(y = 'count',
       x = 'number of runs until relative outer variance convergence')

# plot for grass
p1_grass <- numruns %>%
  ggplot(mapping = aes(x = run, y = relative_outer_var_grass,
                       group = combination_number)) +
  geom_line(alpha = 0.2) +
  geom_hline(yintercept = 0.2, colour = 'darkgreen') +
  labs(y = 'relative outer variance grass',
       title = 'grass') +
  facet_grid(~ tick) +
  theme(axis.title.x = NULL)

p2_grass <- omega_conv_grass %>%
  ggplot(mapping = aes(x = run)) +
  geom_histogram(binwidth = 1) +
  geom_vline(xintercept = mean(omega_conv_grass$run),
             colour = 'darkgreen', lty = 2) +
  geom_vline(xintercept = median(omega_conv_grass$run),
             colour = 'darkblue', lty = 2) +
  annotate(geom = 'text', label = 'mean',
           x = mean(omega_conv_grass$run) + 1, y = 6,
           colour = 'darkgreen', angle = 90) +
  annotate(geom = 'text', label = 'median',
           x = median(omega_conv_grass$run) - 1, y = 6,
           colour = 'darkblue', angle = 90) +
  labs(y = 'count') +
  theme(axis.title.x = NULL)

# plot all together
outer_var_conv_plot <- (p1_sheep + p1_wolves + p1_grass) /
  (p2_sheep + p2_wolves + p2_grass) +
  plot_layout(guides = 'collect') +
  plot_annotation(tag_levels = '1') &
  theme_minimal() +
  theme(plot.tag = element_text(size = 8),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(fill = NA, colour = 'lightgrey'))

outer_var_conv_plot[[1]][[1]] <- outer_var_conv_plot[[1]][[1]] +
  theme(axis.title.x = element_blank())
outer_var_conv_plot[[1]][[2]] <- outer_var_conv_plot[[1]][[2]] +
  theme(axis.title.y = element_blank())
outer_var_conv_plot[[1]][[3]] <- outer_var_conv_plot[[1]][[3]] +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())

outer_var_conv_plot[[2]][[1]] <- outer_var_conv_plot[[2]][[1]] +
  theme(axis.title.x = element_blank())
outer_var_conv_plot[[2]][[2]] <- outer_var_conv_plot[[2]][[2]] +
  theme(axis.title.y = element_blank())
outer_var_conv_plot[[2]][[3]] <- outer_var_conv_plot[[2]][[3]] +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())
```

Inspect all:
```{r}
# mean convergence
mean_conv_plot
ggsave('figures/wolfsheep_mean_convergence.pdf',
       height = 15, width = 20, units = 'cm')

# coefficient of variation
cv_conv_plot
ggsave('figures/wolfsheep_coefficientofvariation.pdf',
       height = 15, width = 20, units = 'cm')

# windowed variance
outer_var_conv_plot
ggsave('figures/wolfsheep_windowed_variance.pdf',
       height = 15, width = 20, units = 'cm')
```
Mean convergence means and medians all under 15.
Coefficient of variation means and medians all under 5.
Outer relative variance means and medians all under 15.

I think I can safely go for 10.

# Sampling parameter space
Sample 1250, 5000, and 10000 samples using first random sampling and then Latin Hypercube Sampling.
There are 7 parameters in this model. The parameter ranges are
- initial-number-sheep 0 - 250
- initial-number-wolves 0 - 250
- grass-regrowth-time 0 - 100
- sheep-gain-from-food 0 - 50
- wolf-grain-from-food 0 - 100
- sheep-reproduce 1 - 20
- wolf-reproduce 0 - 20.
There are 251 * 251 * 101 * 51 * 101 * 20 * 21 = 1.376606e+13 parameter combinations.

I'm going to use 1250, 5000, and 10000 parameter combinations. That constitutes 9.080303e-09, 3.632121e-08, and 7.264243e-08 % of the parameter space, respectively.

Random samples:
```{r}
set.seed(134)

random_sample_ws <- function(size = 1250){
  rand_sample <- tibble(initial_number_sheep = numeric(size),
                        initial_number_wolves = numeric(size),
                        grass_regrowth_time = numeric(size),
                        sheep_gain_from_food = numeric(size),
                        wolf_gain_from_food = numeric(size),
                        sheep_reproduce = numeric(size),
                        wolf_reproduce = numeric(size))
  
  rand_sample$initial_number_sheep <- sample(0:250,
                                             size = size,
                                             replace = TRUE)
  rand_sample$initial_number_wolves <- sample(0:250,
                                              size = size,
                                              replace = TRUE)
  rand_sample$grass_regrowth_time <- sample(0:100,
                                            size = size,
                                            replace = TRUE)
  rand_sample$sheep_gain_from_food <- sample(0:50,
                                             size = size,
                                             replace = TRUE)
  rand_sample$wolf_gain_from_food <- sample(0:100,
                                            size = size,
                                            replace = TRUE)
  rand_sample$sheep_reproduce <- sample(1:20,
                                        size = size,
                                        replace = TRUE)
  rand_sample$wolf_reproduce <- sample(0:20,
                                       size = size,
                                       replace = TRUE)
  
  rand_sample <- rand_sample %>%
    mutate(technique = 'random',
           sample_size = size) %>%
    distinct()
  
  return(rand_sample)
}

# 1250 random samples
rand_sample_1250 <- random_sample_ws(size = 1250)
# check that they are all unique
rand_sample_1250 %>%
  distinct() %>%
  summarise(n())

# add 3750 cases to make up the 5000 sample
rand_sample_5000 <- random_sample_ws(size = 3750)
rand_sample_5000 <- rbind(rand_sample_5000,
                          rand_sample_1250)
rand_sample_5000 <- rand_sample_5000 %>%
  mutate(sample_size = 5000)
# check that they are all unique
rand_sample_5000 %>%
  distinct() %>%
  summarise(n())

# add 5000 cases to make up the 10000 sample
rand_sample_10000 <- random_sample_ws(size = 5000)
rand_sample_10000 <- rbind(rand_sample_10000,
                          rand_sample_5000)
rand_sample_10000 <- rand_sample_10000 %>%
  mutate(sample_size = 10000)
# check that they are all unique
rand_sample_10000 %>%
  distinct() %>%
  summarise(n())
```

Latin Hypercube sampling:
```{r}
lhs_sample_ws <- function(size = 1250){
  # n: number of samples to draw
  # k: number of variables over which to sample
  
  # LHS with n samples over k variables
  A <- randomLHS(n = size, k = 7)
  
  # transform samples to represent the parameter space
  # create empty matrix to hold parameter space samples
  B <- matrix(nrow = nrow(A), ncol = ncol(A))
  # round uniform distribution between 0 and 25 for
  # initial number of sheep and wolves
  B[,1] <- round(qunif(A[,1], min = 0, max = 250), digits = 0)
  B[,2] <- round(qunif(A[,2], min = 0, max = 250), digits = 0)
  # round uniform distribution between 0 and 50 for sheep gain from food
  B[,3] <- round(qunif(A[,3], min = 0, max = 50), digits = 0)
  # round uniform distribution between 0 and 100 for wolf gain from food
  B[,4] <- round(qunif(A[,4], min = 0, max = 100), digits = 0)
  # round uniform distribution between 1 and 20 for 
  # sheep reproduction probability and wolf reproduction probability
  B[,5] <- round(qunif(A[,5], min = 1, max = 20), digits = 0)
  B[,6] <- round(qunif(A[,6], min = 0, max = 20), digits = 0)
  # round uniform distribution between 0 and 100 for grass regrowth time
  B[,7] <- round(qunif(A[,7], min = 0, max = 100), digits = 0)

  # set column names
  colnames(B) <- c('initial_number_sheep', 'initial_number_wolves',
                        'sheep_gain_from_food', 'wolf_gain_from_food',
                        'sheep_reproduce', 'wolf_reproduce',
                        'grass_regrowth_time')
  
  B <- as_tibble(B) %>%
    mutate(technique = 'LHS',
           sample_size = size)
  
  return(B)
}

lhs_sample_1250 <- lhs_sample_ws(size = 1250)
lhs_sample_5000 <- lhs_sample_ws(size = 5000)
lhs_sample_10000 <- lhs_sample_ws(size = 10000)
```

Compile into 1 table and save to import in Python
```{r}
parameter_space_samples <- bind_rows(rand_sample_1250,
                                     rand_sample_5000,
                                     rand_sample_10000,
                                     lhs_sample_1250,
                                     lhs_sample_5000,
                                     lhs_sample_10000)

write_csv(as.data.frame(parameter_space_samples),
          'data/processed/wolfsheep_parameter_sample.csv')
```

For now, run this:
```{r}
parameter_space_samples %>%
  filter(sample_size == 1250,
         technique == 'random') %>%
  write_csv('data/processed/wolfsheep_parameter_sample_1.csv')
```

Run each parameter combination 15 times.
```{python}
all_comb = pd.read_csv('C:\\Users\\Femke Keij\\OneDrive\\Thesis\\R projects\\LU thesis GitHub\\data\\processed\\wolfsheep_parameter_sample_1.csv')

parameter_space_df = pd.DataFrame(columns = ['technique', 'sample_size', 'run', 'tick' ,'initial_number_sheep', 'initial_number_wolves', 'sheep_gain_from_food', 'wolf_gain_from_food', 'sheep_reproduce', 'wolf_reproduce', 'grass_regrowth', 'sheep', 'wolves', 'grass'])

# RUNNING IN GROUPS OF 10
all_comb = all_comb.iloc[321:341, 0:9]
all_comb = all_comb.reset_index()
# next: 281:301
# next: 301:321
# next: 321:341
# next: 341:361

# set random seed
#netlogo.command('clean')

# for each combination of parameters
for i in range(0, len(all_comb)):
  # retrieve parameters for run
  initial_number_sheep = str(all_comb['initial_number_sheep'][i])
  initial_number_wolves = str(all_comb['initial_number_wolves'][i])
  sheep_gain_from_food = str(all_comb['sheep_gain_from_food'][i])
  wolf_gain_from_food = str(all_comb['wolf_gain_from_food'][i])
  sheep_reproduce = str(all_comb['sheep_reproduce'][i])
  wolf_reproduce = str(all_comb['wolf_reproduce'][i])
  grass_regrowth = str(all_comb['grass_regrowth_time'][i])
  # retrieve extra info
  technique = str(all_comb['technique'][i])
  sample_size = str(all_comb['sample_size'][i])
  
  # set parameters
  netlogo.command(''.join(['set initial-number-sheep ', initial_number_sheep]))
  netlogo.command(''.join(['set initial-number-wolves ', initial_number_wolves]))
  netlogo.command(''.join(['set sheep-gain-from-food ', sheep_gain_from_food]))
  netlogo.command(''.join(['set wolf-gain-from-food ', wolf_gain_from_food]))
  netlogo.command(''.join(['set sheep-reproduce ', sheep_reproduce]))
  netlogo.command(''.join(['set wolf-reproduce ', wolf_reproduce]))
  netlogo.command(''.join(['set grass-regrowth-time ', grass_regrowth]))
  
  # run combination 10 times
  for j in range(1, 16):
    # set up
    netlogo.command('setup')
    # execute 1 tick at a time, and report
    for h in range(0, 500):
      netlogo.command('go')
      # record number of sheep, wolves, grass, and ticks
      sheep = netlogo.report('count sheep')
      wolves = netlogo.report('count wolves')
      grass = netlogo.report('count grass')
      ticks = netlogo.report('ticks')
      # add to data frame
      parameter_space_df = parameter_space_df.append({'technique': technique, 'sample_size': sample_size, 'run': j, 'tick': ticks, 'initial_number_sheep': initial_number_sheep, 'initial_number_wolves': initial_number_wolves, 'sheep_gain_from_food': sheep_gain_from_food, 'wolf_gain_from_food': wolf_gain_from_food, 'sheep_reproduce': sheep_reproduce, 'wolf_reproduce': wolf_reproduce, 'grass_regrowth': grass_regrowth, 'sheep': sheep, 'wolves': wolves, 'grass': grass}, ignore_index = True)
      # break out of this level if the maximum number of sheep is reached and there are no wolves or grass left
      if wolves == 0 and grass == 0 and sheep > 30000:
        break
      
  print(''.join(['finished run ', str(i), '\n']), flush = True)

# WHEN RUNNING IN GROUPS, CHANGE SAVE NUMBER
# save data as csv file
parameter_space_df.to_csv('C:\\Users\\Femke Keij\\OneDrive\\Thesis\\R projects\\LU thesis GitHub\\data\\raw\\wolfsheep_output18.csv')
```

Close the model:
```{python}
netlogo.kill_workspace()
```

Stop running the python environment:
```{python}
quit
```

# Summarise repeated runs
Clear working directory:
```{r}
rm(list = ls(all = TRUE))
```

Import csv output format for the model as tidy data. We ignore the first column because it contains the indices.
This is a limited sample data set with around 300 cases that will be used for now.
```{r}
ws_output <- read_csv("data/raw/wolfsheep_output200.csv",
                        col_names = TRUE)[, - 1]
```

Look at the distribution of the outputs within 1 parameter combination. If this is approximately normal (or concentrated around a single mean), we can summarise the data into a single value (the mean).
```{r}
ws_output <- ws_output %>%
  group_by(initial_number_sheep, initial_number_wolves,
           sheep_gain_from_food, wolf_gain_from_food,
           sheep_reproduce, wolf_reproduce,
           grass_regrowth) %>%
  mutate(combination_number = cur_group_id()) %>%
  filter(combination_number <= 200)

# plot a histogram of all the standard deviations for the sheep output
sd_plot_sheep <- ws_output %>%
  filter(tick %in% c(50, 250, 450)) %>%
  group_by(combination_number, tick) %>%
  mutate(sd = sd(sheep)) %>%
  select(combination_number, sd, tick) %>%
  distinct() %>%
  drop_na() %>%
  ggplot(mapping = aes(x = sd)) +
  geom_histogram(binwidth = 100) +
  facet_grid(~ tick) +
  theme_minimal() +
  theme(panel.border = element_rect(fill = NA,
                                    colour = 'lightgrey'),
        panel.grid.minor = element_blank()) +
  labs(x = 'sd of number of sheep per parameter combination')

sd_plot_wolves <- ws_output %>%
  filter(tick %in% c(50, 250, 450)) %>%
  group_by(combination_number, tick) %>%
  mutate(sd = sd(wolves)) %>%
  select(combination_number, sd, tick) %>%
  distinct() %>%
  drop_na() %>%
  ggplot(mapping = aes(x = sd)) +
  geom_histogram(binwidth = 100) +
  facet_grid(~ tick) +
  theme_minimal() +
  theme(panel.border = element_rect(fill = NA,
                                    colour = 'lightgrey'),
        panel.grid.minor = element_blank()) +
  labs(x = 'sd of number of wolves per parameter combination')

sd_plot_grass <- ws_output %>%
  filter(tick %in% c(50, 250, 450)) %>%
  group_by(combination_number, tick) %>%
  mutate(sd = sd(grass)) %>%
  select(combination_number, sd, tick) %>%
  distinct() %>%
  drop_na() %>%
  ggplot(mapping = aes(x = sd)) +
  geom_histogram(binwidth = 100) +
  facet_grid(~ tick) +
  theme_minimal() +
  theme(panel.border = element_rect(fill = NA,
                                    colour = 'lightgrey'),
        panel.grid.minor = element_blank()) +
  labs(x = 'sd of amount of gras per parameter combination')

# inspect the parameter combinations that yield large standard deviations
large_sd <- ws_output %>%
  filter(tick %in% c(50, 250, 450)) %>%
  group_by(combination_number, tick) %>%
  mutate(sd_sheep = sd(sheep),
         sd_wolves = sd(wolves),
         sd_grass = sd(grass),
         mean_sheep = mean(sheep),
         mean_wolves = mean(wolves),
         mean_grass = mean(grass)) %>%
  select(- c(run, technique, sample_size)) %>%
  distinct() %>%
  ungroup()

# show the distributions for the parameter combinations with the 5
# largest standard deviations
large_sd_sheep <- large_sd %>%
  select(- c(sheep, wolves, grass)) %>%
  distinct() %>%
  arrange(- sd_sheep) %>%
  mutate(rank_sd = row_number()) %>%
  filter(rank_sd <= 5) %>%
  select(combination_number)
large_sd_wolves <- large_sd %>%
  select(- c(sheep, wolves, grass)) %>%
  distinct() %>%
  arrange(- sd_wolves) %>%
  mutate(rank_sd = row_number()) %>%
  filter(rank_sd <= 5) %>%
  select(combination_number)
large_sd_grass <- large_sd %>%
  select(- c(sheep, wolves, grass)) %>%
  distinct() %>%
  arrange(- sd_grass) %>%
  mutate(rank_sd = row_number()) %>%
  filter(rank_sd <= 5) %>%
  select(combination_number)

individual_plot_sheep <- large_sd %>%
  filter(combination_number %in%
           unlist(large_sd_sheep)) %>%
  ggplot(mapping = aes(x = sheep)) +
  geom_histogram(binwidth = 100) +
  geom_vline(aes(xintercept = mean_sheep,
                 group = combination_number),
             colour = 'red') +
  facet_grid(combination_number ~ tick,
             scales = 'free') +
  labs(x = 'amount of sheep')

individual_plot_wolves <- large_sd %>%
  filter(combination_number %in%
           unlist(large_sd_wolves)) %>%
  ggplot(mapping = aes(x = wolves)) +
  geom_histogram(binwidth = 100) +
  geom_vline(aes(xintercept = mean_wolves,
                 group = combination_number),
             colour = 'red') +
  facet_grid(combination_number ~ tick,
             scales = 'free') +
  labs(x = 'amount of wolves')

individual_plot_grass <- large_sd %>%
  filter(combination_number %in%
           unlist(large_sd_grass)) %>%
  ggplot(mapping = aes(x = grass)) +
  geom_histogram(binwidth = 100) +
  geom_vline(aes(xintercept = mean_grass,
                 group = combination_number),
             colour = 'red') +
  facet_grid(combination_number ~ tick,
             scales = 'free') +
  labs(x = 'amount of grass')

# combined plot for sheep
sd_plot_sheep / individual_plot_sheep +
  plot_layout(guides = 'collect') +
  plot_annotation(tag_levels = 'A') &
  theme_minimal() +
  theme(plot.tag = element_text(size = 8),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(fill = NA, colour = 'lightgrey'))

ggsave('figures/wolfsheep_summarising_sheep.pdf',
       height = 30, width = 20, units = 'cm')

# combined plot for wolves
sd_plot_wolves / individual_plot_wolves +
  plot_layout(guides = 'collect') +
  plot_annotation(tag_levels = 'A') &
  theme_minimal() +
  theme(plot.tag = element_text(size = 8),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(fill = NA, colour = 'lightgrey'))

ggsave('figures/wolfsheep_summarising_wolves.pdf',
       height = 30, width = 20, units = 'cm')

# combined plot for grass
sd_plot_grass / individual_plot_grass +
  plot_layout(guides = 'collect') +
  plot_annotation(tag_levels = 'A') &
  theme_minimal() +
  theme(plot.tag = element_text(size = 8),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(fill = NA, colour = 'lightgrey'))

ggsave('figures/wolfsheep_summarising_grass.pdf',
       height = 30, width = 20, units = 'cm')
```

Manipulate data so that there is 1 outcome for each parameter combination:
The outcomes for each parameter set are summarized by their mean sheep, wolves, and grass at every tick of the simulation.
```{r}
# group the data by parameter combination and ticks
ws_summary <- ws_output %>%
  group_by(combination_number, tick) %>%
  mutate(mean_sheep = mean(sheep),
         mean_wolves = mean(wolves),
         mean_grass = mean(grass)) %>%
  mutate(sample_size = 200) %>%
  select(- c(run, sheep, wolves, grass)) %>%
  distinct() %>%
  rename(wolves_reproduce = wolf_reproduce,
         wolves_gain_from_food = wolf_gain_from_food,
         ticks = tick,
         sheep = mean_sheep,
         wolves = mean_wolves,
         grass = mean_grass)

ws_summary %>%
  write_csv(file = 'data/processed/wolfsheep_summary200.csv')
```

Also save non-summarised version:
```{r}
ws_output %>%
  mutate(sample_size = 200) %>%
  rename(wolves_reproduce = wolf_reproduce,
         wolves_gain_from_food = wolf_gain_from_food,
         ticks = tick) %>%
  write_csv(file = 'data/processed/wolfsheep_full200.csv')
```