---
title: "WolfSheep_00_Run_Model"
author: "Femke Keij S2647168"
date: "2023-03-13"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import Python modules:
```{python}
import pandas as pd
import matplotlib.pyplot as plt
import seaborn as sns
import math
import statistics
import itertools
sns.set_style('white')
sns.set_context('talk')

import pyNetLogo # to run NetLogo from RStudio
```

Import R packages:
```{r}
# for importing / working with tidy data
library(tidyverse)
# for arranging plots
library(patchwork)
# for cumulative computations
library(zoo)
# for rolling averages
library(slider)
```

Set seed:
```{r}
set.seed(42)
```

# Running all possible parameter combinations
Link to and start the Fire model:
```{python}
netlogo = pyNetLogo.NetLogoLink(gui = True) # start netlogo

netlogo.load_model(r"C:\Users\Femke Keij\OneDrive\Thesis\R projects\LU thesis local\models\WolfSheep_myversion.nlogo")
# open model

# set to correct model version
netlogo.command('set model-version "sheep-wolves-grass"')
```

# Determine number of runs
1) coefficient of variation
2) mean convergence
3) windowed variance

then also try 1) and 2) with setting a different epsilon for each parameter (only necessary if they aren't on the same scale, so maybe different for grass vs. wolf and sheep? check this before trying it out)

We can compute all these statistics using just 1 data set. To do so, we run each parameter combination 50 times and save the outputs at time step 100 (this is an arbitrary choice that should potentially be remedied later to include the time steps etc. that I'm actually going to use). I then compute the following:
1) coefficient of variation: sigma / mu for 1, ..., 50 runs
2) mean convergence: mu for 1, ..., 50 runs
3) windowed variance: sigma for 1, ..., 50 runs
We then see where these stabilize within a given epsilon.

Instead of combining a low, mid, and high value for each parameter, I use only the 25% and 75% quartile of each range. This brings the number of parameter combinations down from 2187 to 128.
```{python}
initial_number_sheep = [62, 188]
initial_number_wolves = [62, 188]
sheep_gain_from_food = [12, 38]
wolf_gain_from_food = [25, 72]
sheep_reproduce = [5, 15]
wolf_reproduce = [5, 15]
grass_regrowth = [25, 72]

dictionary = {'initial_number_sheep': [62, 188], 'initial_number_wolves': [62, 188], 'sheep_gain_from_food': [12, 38], 'wolf_gain_from_food': [25, 72], 'sheep_reproduce': [5, 15], 'wolf_reproduce': [5, 15], 'grass_regrowth': [25, 72]}

def expand_grid(data_dict):
    rows = itertools.product(*data_dict.values())
    return pd.DataFrame.from_records(rows, columns = data_dict.keys())

param_comb = expand_grid(dictionary)
```

Obtain all necessary data:
```{python}
# data frame to store results
out_df = pd.DataFrame(columns = ['initial_number_sheep', 'initial_number_wolves', 'sheep_gain_from_food', 'wolf_gain_from_food', 'sheep_reproduce', 'wolf_reproduce', 'grass_regrowth', 'run', 'wolves', 'sheep', 'grass'])
# add empty row to start
out_df = out_df.append({'initial_number_sheep': 0, 'initial_number_wolves': 0, 'sheep_gain_from_food': 0, 'wolf_gain_from_food': 0, 'sheep_reproduce': 0, 'wolf_reproduce': 0, 'grass_regrowth': 0, 'run': 0, 'sheep': -100, 'wolves': -100, 'grass': -100}, ignore_index = True)

# set counter
count = 0
# set random seed
netlogo.command('clean')

for i in range(0,len(param_comb)):
  # set parameters
  initial_number_sheep = str(param_comb['initial_number_sheep'][i])
  initial_number_wolves = str(param_comb['initial_number_wolves'][i])
  sheep_gain_from_food = str(param_comb['sheep_gain_from_food'][i])
  wolf_gain_from_food = str(param_comb['wolf_gain_from_food'][i])
  sheep_reproduce = str(param_comb['sheep_reproduce'][i])
  wolf_reproduce = str(param_comb['wolf_reproduce'][i])
  grass_regrowth = str(param_comb['grass_regrowth'][i])
  netlogo.command(''.join(['set initial-number-sheep ', initial_number_sheep]))
  netlogo.command(''.join(['set initial-number-wolves ', initial_number_wolves]))
  netlogo.command(''.join(['set  sheep-gain-from-food ',  sheep_gain_from_food]))
  netlogo.command(''.join(['set wolf-gain-from-food ',  wolf_gain_from_food]))
  netlogo.command(''.join(['set sheep-reproduce ', sheep_reproduce]))
  netlogo.command(''.join(['set wolf-reproduce ', wolf_reproduce]))
  netlogo.command(''.join(['set grass-regrowth-time ', grass_regrowth]))
  
  for x in range(0, 50):
    # setup
    netlogo.command('setup')
    # execute for 100 ticks
    netlogo.command('repeat 100 [go]')
    # record sheep, wolves, grass
    sheep = netlogo.report('count sheep')
    wolves = netlogo.report('count wolves')
    grass = netlogo.report ('count grass')
    # store
    out_df = out_df.append({'initial_number_sheep': initial_number_sheep, 'initial_number_wolves': initial_number_wolves, 'sheep_gain_from_food': sheep_gain_from_food, 'wolf_gain_from_food': wolf_gain_from_food, 'sheep_reproduce': sheep_reproduce, 'wolf_reproduce': wolf_reproduce, 'grass_regrowth': grass_regrowth, 'run': x + 1, 'sheep': sheep, 'wolves': wolves, 'grass': grass}, ignore_index = True)

out_df.to_csv('data/raw/wolfsheep_numruns.csv')
```

Compute statistics:
```{r}
numruns <- read.csv('data/raw/wolfsheep_numruns.csv')

# remove first row
numruns <- numruns %>%
  slice(-1)

# add a unique number for each parameter combination
numruns <- numruns %>%
  group_by(initial_number_sheep, initial_number_wolves, sheep_gain_from_food,
           wolf_gain_from_food, sheep_reproduce, wolf_reproduce, grass_regrowth) %>%
  mutate(combination_number = cur_group_id())
```

1) Mean convergence
With $\epsilon_\mu = 1$
```{r}
numruns <- numruns %>%
  group_by(combination_number) %>%
  # compute cumulative means
  # and differences between means
  mutate(mean_sheep = cummean(sheep),
         mean_wolves = cummean(wolves),
         mean_grass = cummean(grass),
         mean_diff_sheep = mean_sheep - lag(mean_sheep, default = mean_sheep[1]),
         mean_diff_wolves = mean_wolves - lag(mean_wolves, default = mean_wolves[1]),
         mean_diff_grass = mean_grass - lag(mean_grass, default = mean_grass[1]))

# compute at which point the mean difference settles below 1
mean_conv_sheep <- numruns %>%
  group_by(combination_number) %>%
  filter(abs(mean_diff_sheep) < 1,
         run != 1) %>%
  slice(which.min(run))

mean_conv_wolves <- numruns %>%
  group_by(combination_number) %>%
  filter(abs(mean_diff_wolves) < 1,
         run != 1) %>%
  slice(which.min(run))

mean_conv_grass <- numruns %>%
  group_by(combination_number) %>%
  filter(abs(mean_diff_grass) < 1,
         run != 1) %>%
  slice(which.min(run))
  
# plot for sheep
p1_sheep <- numruns %>%
  ggplot(mapping = aes(x = run, y = mean_diff_sheep,
                       group = combination_number)) +
  geom_line(colour = 'lightgrey') +
  geom_hline(yintercept = 1, colour = 'darkgreen') +
  geom_hline(yintercept = - 1, colour = 'darkgreen') +
  theme_minimal() +
  labs(y = 'difference mean number of sheep',
       caption = 'Difference in mean number of sheep when adding 1 run')

p2_sheep <- mean_conv_sheep %>%
  ggplot(mapping = aes(x = run)) +
  geom_histogram(binwidth = 1) +
  geom_vline(xintercept = mean(mean_conv_sheep$run),
             colour = 'darkgreen', lty = 2) +
  geom_vline(xintercept = median(mean_conv_sheep$run),
             colour = 'darkblue', lty = 2) +
  annotate(geom = 'text', label = 'mean',
           x = mean(mean_conv_sheep$run) + 1, y = 20,
           colour = 'darkgreen', angle = 90) +
  annotate(geom = 'text', label = 'median',
           x = median(mean_conv_sheep$run) - 1, y = 20,
           colour = 'darkblue', angle = 90) +
  theme_minimal() +
  labs(y = 'count',
       x = 'number of runs until mean convergence')

# plot for wolves
p1_wolves <- numruns %>%
  ggplot(mapping = aes(x = run, y = mean_diff_wolves,
                       group = combination_number)) +
  geom_line(colour = 'lightgrey') +
  geom_hline(yintercept = 1, colour = 'darkgreen') +
  geom_hline(yintercept = - 1, colour = 'darkgreen') +
  theme_minimal() +
  labs(y = 'difference mean number of wolves',
       caption = 'Difference in mean number of wolves when adding 1 run')

p2_wolves <- mean_conv_wolves %>%
  ggplot(mapping = aes(x = run)) +
  geom_histogram(binwidth = 1) +
  geom_vline(xintercept = mean(mean_conv_wolves$run),
             colour = 'darkgreen', lty = 2) +
  geom_vline(xintercept = median(mean_conv_wolves$run),
             colour = 'darkblue', lty = 2) +
  annotate(geom = 'text', label = 'mean',
           x = mean(mean_conv_wolves$run) + 1, y = 20,
           colour = 'darkgreen', angle = 90) +
  annotate(geom = 'text', label = 'median',
           x = median(mean_conv_wolves$run) - 1, y = 20,
           colour = 'darkblue', angle = 90) +
  theme_minimal() +
  labs(y = 'count',
       x = 'number of runs until mean convergence')

# plot for grass
p1_grass <- numruns %>%
  ggplot(mapping = aes(x = run, y = mean_diff_grass,
                       group = combination_number)) +
  geom_line(colour = 'lightgrey') +
  geom_hline(yintercept = 1, colour = 'darkgreen') +
  geom_hline(yintercept = - 1, colour = 'darkgreen') +
  theme_minimal() +
  labs(y = 'difference mean number of grass',
       caption = 'Difference in mean number of grass when adding 1 run')

p2_grass <- mean_conv_grass %>%
  ggplot(mapping = aes(x = run)) +
  geom_histogram(binwidth = 1) +
  geom_vline(xintercept = mean(mean_conv_grass$run),
             colour = 'darkgreen', lty = 2) +
  geom_vline(xintercept = median(mean_conv_grass$run),
             colour = 'darkblue', lty = 2) +
  annotate(geom = 'text', label = 'mean',
           x = mean(mean_conv_grass$run) + 1, y = 20,
           colour = 'darkgreen', angle = 90) +
  annotate(geom = 'text', label = 'median',
           x = median(mean_conv_grass$run) - 1, y = 20,
           colour = 'darkblue', angle = 90) +
  theme_minimal() +
  labs(y = 'count',
       x = 'number of runs until mean convergence')

# plot all together
mean_conv_plot <- (p1_sheep + p1_wolves + p1_grass) / (p2_sheep + p2_wolves + p2_grass)
```

2) Coefficient of variation
With $\epsilon_{c_\upsilon} = 0.1$.
```{r}
numruns <- numruns %>%
  group_by(combination_number) %>%
  # compute cumulative standard deviations
  # compute coefficient of variation
  mutate(sd_sheep = rollapplyr(sheep, 1:n(), sd),
         sd_wolves = rollapplyr(wolves, 1:n(), sd),
         sd_grass = rollapplyr(grass, 1:n(), sd),
         cv_sheep = sd_sheep / mean_sheep,
         cv_wolves = sd_wolves / mean_wolves,
         cv_grass = sd_grass / mean_grass) %>%
  # replace all NAs (resulting from sd = 0) with 0
  replace(is.na(.), 0) %>%
  # compute differences in cvs
  mutate(cv_diff_sheep = cv_sheep - lag(cv_sheep, default = cv_sheep[1]),
         cv_diff_wolves = cv_wolves - lag(cv_wolves, default = cv_wolves[1]),
         cv_diff_grass = cv_grass - lag(cv_grass, default = cv_grass[1]))
  
# compute at which point the mean difference settles below 1
cv_conv_sheep <- numruns %>%
  group_by(combination_number) %>%
  filter(abs(cv_diff_sheep) < 0.1,
         run != 1) %>%
  slice(which.min(run))

cv_conv_wolves <- numruns %>%
  group_by(combination_number) %>%
  filter(abs(cv_diff_wolves) < 0.1,
         run != 1) %>%
  slice(which.min(run))

cv_conv_grass <- numruns %>%
  group_by(combination_number) %>%
  filter(abs(cv_diff_grass) < 0.1,
         run != 1) %>%
  slice(which.min(run))

# plot for sheep
p1_sheep <- numruns %>%
  ggplot(mapping = aes(x = run, y = cv_diff_sheep,
                       group = combination_number)) +
  geom_line(colour = 'lightgrey') +
  geom_hline(yintercept = 0.1, colour = 'darkgreen') +
  geom_hline(yintercept = - 0.1, colour = 'darkgreen') +
  theme_minimal() +
  labs(y = 'cv number of sheep',
       caption = 'Difference in cv of sheep when adding 1 run')

p2_sheep <- cv_conv_sheep %>%
  ggplot(mapping = aes(x = run)) +
  geom_histogram(binwidth = 1) +
  geom_vline(xintercept = mean(cv_conv_sheep$run),
             colour = 'darkgreen', lty = 2) +
  geom_vline(xintercept = median(cv_conv_sheep$run),
             colour = 'darkblue', lty = 2) +
  annotate(geom = 'text', label = 'mean',
           x = mean(cv_conv_sheep$run) + 1, y = 20,
           colour = 'darkgreen', angle = 90) +
  annotate(geom = 'text', label = 'median',
           x = median(cv_conv_sheep$run) - 1, y = 20,
           colour = 'darkblue', angle = 90) +
  theme_minimal() +
  labs(y = 'count',
       x = 'number of runs until cv convergence')

# plot for wolves
p1_wolves <- numruns %>%
  ggplot(mapping = aes(x = run, y = cv_diff_wolves,
                       group = combination_number)) +
  geom_line(colour = 'lightgrey') +
  geom_hline(yintercept = 0.1, colour = 'darkgreen') +
  geom_hline(yintercept = - 0.1, colour = 'darkgreen') +
  theme_minimal() +
  labs(y = 'difference cv of wolves',
       caption = 'Difference in cv of wolves when adding 1 run')

p2_wolves <- cv_conv_wolves %>%
  ggplot(mapping = aes(x = run)) +
  geom_histogram(binwidth = 1) +
  geom_vline(xintercept = mean(cv_conv_wolves$run),
             colour = 'darkgreen', lty = 2) +
  geom_vline(xintercept = median(cv_conv_wolves$run),
             colour = 'darkblue', lty = 2) +
  annotate(geom = 'text', label = 'mean',
           x = mean(cv_conv_wolves$run) + 1, y = 20,
           colour = 'darkgreen', angle = 90) +
  annotate(geom = 'text', label = 'median',
           x = median(cv_conv_wolves$run) - 1, y = 20,
           colour = 'darkblue', angle = 90) +
  theme_minimal() +
  labs(y = 'count',
       x = 'number of runs until cv convergence')

# plot for grass
p1_grass <- numruns %>%
  ggplot(mapping = aes(x = run, y = cv_diff_grass,
                       group = combination_number)) +
  geom_line(colour = 'lightgrey') +
  geom_hline(yintercept = 0.1, colour = 'darkgreen') +
  geom_hline(yintercept = - 0.1, colour = 'darkgreen') +
  theme_minimal() +
  labs(y = 'difference cv of grass',
       caption = 'Difference in cv of grass when adding 1 run')

p2_grass <- cv_conv_grass %>%
  ggplot(mapping = aes(x = run)) +
  geom_histogram(binwidth = 1) +
  geom_vline(xintercept = mean(cv_conv_grass$run),
             colour = 'darkgreen', lty = 2) +
  geom_vline(xintercept = median(cv_conv_grass$run),
             colour = 'darkblue', lty = 2) +
  annotate(geom = 'text', label = 'mean',
           x = mean(cv_conv_grass$run) + 1, y = 20,
           colour = 'darkgreen', angle = 90) +
  annotate(geom = 'text', label = 'median',
           x = median(cv_conv_grass$run) - 1, y = 20,
           colour = 'darkblue', angle = 90) +
  theme_minimal() +
  labs(y = 'count',
       x = 'number of runs until cv convergence')

# plot all together
cv_conv_plot <- (p1_sheep + p1_wolves + p1_grass) / (p2_sheep + p2_wolves + p2_grass)
```

3) Windowed variance / relative outer variance
```{r}
numruns <- numruns %>% 
  group_by(combination_number) %>% 
  mutate(windowed_var_sheep = slider::slide_dbl(sheep, var, .before = 5, .after = 5),
         windowed_var_wolves = slider::slide_dbl(wolves, var, .before = 5, .after = 5),
         windowed_var_grass = slider::slide_dbl(grass, var, .before = 5, .after = 5),
         relative_outer_var_sheep = windowed_var_sheep / max(windowed_var_sheep),
         relative_outer_var_wolves = windowed_var_wolves / max(windowed_var_wolves),
         relative_outer_var_grass = windowed_var_grass / max(windowed_var_grass)) %>%
  # there's a few instances where the relative outer variance is NaN because the population
  # is 0, therefore, we set those to 0 instead of NaN
  mutate_at(vars(relative_outer_var_sheep,
                 relative_outer_var_grass,
                 relative_outer_var_wolves), ~ replace(., is.nan(.), 0))

# compute at which point the relative outer variance settles below 0.2
omega_conv_sheep <- numruns %>%
  group_by(combination_number) %>%
  filter(abs(relative_outer_var_sheep) < 0.2,
         run != 1) %>%
  slice(which.min(run))

omega_conv_wolves <- numruns %>%
  group_by(combination_number) %>%
  filter(abs(relative_outer_var_wolves) < 0.2,
         run != 1) %>%
  slice(which.min(run))

omega_conv_grass <- numruns %>%
  group_by(combination_number) %>%
  filter(abs(relative_outer_var_grass) < 0.2,
         run != 1) %>%
  slice(which.min(run)) 

# plot for sheep
p1_sheep <- numruns %>%
  ggplot(mapping = aes(x = run, y = relative_outer_var_sheep,
                       group = combination_number)) +
  geom_line(colour = 'lightgrey') +
  geom_hline(yintercept = 0.2, colour = 'darkgreen') +
  labs(y = 'relative outer variance sheep') +
  theme_minimal()

p2_sheep <- omega_conv_sheep %>%
  ggplot(mapping = aes(x = run)) +
  geom_histogram(binwidth = 1) +
  geom_vline(xintercept = mean(omega_conv_sheep$run),
             colour = 'darkgreen', lty = 2) +
  geom_vline(xintercept = median(omega_conv_sheep$run),
             colour = 'darkblue', lty = 2) +
  annotate(geom = 'text', label = 'mean',
           x = mean(omega_conv_sheep$run) + 1, y = 20,
           colour = 'darkgreen', angle = 90) +
  annotate(geom = 'text', label = 'median',
           x = median(omega_conv_sheep$run) - 1, y = 20,
           colour = 'darkblue', angle = 90) +
  theme_minimal() +
  labs(y = 'count',
       x = 'number of runs until relative outer variance convergence')

# plot for wolves
p1_wolves <- numruns %>%
  ggplot(mapping = aes(x = run, y = relative_outer_var_wolves,
                       group = combination_number)) +
  geom_line(colour = 'lightgrey') +
  geom_hline(yintercept = 0.2, colour = 'darkgreen') +
  labs(y = 'relative outer variance wolves') +
  theme_minimal()

p2_wolves <- omega_conv_wolves %>%
  ggplot(mapping = aes(x = run)) +
  geom_histogram(binwidth = 1) +
  geom_vline(xintercept = mean(omega_conv_wolves$run),
             colour = 'darkgreen', lty = 2) +
  geom_vline(xintercept = median(omega_conv_wolves$run),
             colour = 'darkblue', lty = 2) +
  annotate(geom = 'text', label = 'mean',
           x = mean(omega_conv_wolves$run) + 1, y = 20,
           colour = 'darkgreen', angle = 90) +
  annotate(geom = 'text', label = 'median',
           x = median(omega_conv_wolves$run) - 1, y = 20,
           colour = 'darkblue', angle = 90) +
  theme_minimal() +
  labs(y = 'count',
       x = 'number of runs until relative outer variance convergence')

# plot for grass
p1_grass <- numruns %>%
  ggplot(mapping = aes(x = run, y = relative_outer_var_grass,
                       group = combination_number)) +
  geom_line(colour = 'lightgrey') +
  geom_hline(yintercept = 0.2, colour = 'darkgreen') +
  labs(y = 'relative outer variance grass') +
  theme_minimal()

p2_grass <- omega_conv_grass %>%
  ggplot(mapping = aes(x = run)) +
  geom_histogram(binwidth = 1) +
  geom_vline(xintercept = mean(omega_conv_grass$run),
             colour = 'darkgreen', lty = 2) +
  geom_vline(xintercept = median(omega_conv_grass$run),
             colour = 'darkblue', lty = 2) +
  annotate(geom = 'text', label = 'mean',
           x = mean(omega_conv_grass$run) + 1, y = 20,
           colour = 'darkgreen', angle = 90) +
  annotate(geom = 'text', label = 'median',
           x = median(omega_conv_grass$run) - 1, y = 20,
           colour = 'darkblue', angle = 90) +
  theme_minimal() +
  labs(y = 'count',
       x = 'number of runs until relative outer variance convergence')

# plot all together
outer_var_conv_plot <- (p1_sheep + p1_wolves + p1_grass) / (p2_sheep + p2_wolves + p2_grass)
```

Inspect all results:
```{r}
mean_conv_plot
cv_conv_plot
outer_var_conv_plot
```
Mean convergence means and medians all under 15.
Coefficient of variation means and medians all under 5.
Outer relative variance means and medians all under 15.

I think I can safely go for 10.

Before making this conclusion final, check that I've done the outer relative variance thing correctly.

# Sampling parameter space
I set up a Latin Hypercube sample of the parameter space to ensure adequate coverage. Then I run the model 15 times for each parameter combination.
```{r}
# for Latin hypercube sampling
library(lhs)

# n: number of samples to draw
# k: number of variables over which to sample
  
# LHS with n samples over k variables
A <- randomLHS(n = 40, k = 7)
  
# transform samples to represent the parameter space
# create empty matrix to hold parameter space samples
B <- matrix(nrow = nrow(A), ncol = ncol(A))
# round uniform distribution between 0 and 25 for
# initial number of sheep and wolves
B[,1] <- round(qunif(A[,1], min = 0, max = 250), digits = 0)
B[,2] <- round(qunif(A[,2], min = 0, max = 250), digits = 0)
# round uniform distribution between 0 and 50 for sheep gain from food
B[,3] <- round(qunif(A[,3], min = 0, max = 50), digits = 0)
# round uniform distribution between 0 and 100 for wolf gain from food
B[,4] <- round(qunif(A[,4], min = 0, max = 100), digits = 0)
# round uniform distribution between 1 and 20 for 
# sheep reproduction probability and wolf reproduction probability
B[,5] <- round(qunif(A[,5], min = 1, max = 20), digits = 0)
B[,6] <- round(qunif(A[,6], min = 1, max = 20), digits = 0)
# round uniform distribution between 0 and 100 for grass regrowth time
B[,7] <- round(qunif(A[,7], min = 0, max = 100), digits = 0)

# set column names
colnames(B) <- c('initial_number_sheep', 'initial_number_wolves',
                        'sheep_gain_from_food', 'wolf_gain_from_food',
                        'sheep_reproduce', 'wolf_reproduce',
                        'grass_regrowth')
```
Inspect sample to see if I'm happy with it:
```{r}
summary(B)
``` 
Save to import in Python:
```{r}
write_csv(as.data.frame(B), 'data/processed/wolfsheep_parameterspace.csv')
```

```{python}
all_comb = pd.read_csv('data/processed/wolfsheep_parameterspace.csv')

parameter_space_df = pd.DataFrame(columns = ['run', 'initial_number_sheep', 'initial_number_wolves', 'sheep_gain_from_food', 'wolf_gain_from_food', 'sheep_reproduce', 'wolf_reproduce', 'grass_regrowth', 'sheep', 'wolves', 'grass', 'ticks', 'combination_number'])

# set random seed
netlogo.command('clean')

# for each combination of parameters
for i in range(0, len(all_comb)):
  # retrieve parameters for run
  initial_number_sheep = str(all_comb['initial_number_sheep'][i])
  initial_number_wolves = str(all_comb['initial_number_wolves'][i])
  sheep_gain_from_food = str(all_comb['sheep_gain_from_food'][i])
  wolf_gain_from_food = str(all_comb['wolf_gain_from_food'][i])
  sheep_reproduce = str(all_comb['sheep_reproduce'][i])
  wolf_reproduce = str(all_comb['wolf_reproduce'][i])
  grass_regrowth = str(all_comb['grass_regrowth'][i])
  
  # set parameters
  netlogo.command(''.join(['set initial-number-sheep ', initial_number_sheep]))
  netlogo.command(''.join(['set initial-number-wolves ', initial_number_wolves]))
  netlogo.command(''.join(['set sheep-gain-from-food ', sheep_gain_from_food]))
  netlogo.command(''.join(['set wolf-gain-from-food ', wolf_gain_from_food]))
  netlogo.command(''.join(['set sheep-reproduce ', sheep_reproduce]))
  netlogo.command(''.join(['set wolf-reproduce ', wolf_reproduce]))
  netlogo.command(''.join(['set grass-regrowth-time ', grass_regrowth]))
  
  # run combination 15 times
  for j in range(1, 16):
    # set up
    netlogo.command('setup')
    # execute 1 tick at a time, and report
    for h in range(0, 500):
      netlogo.command('go')
      # record number of sheep, wolves, grass, and ticks
      sheep = netlogo.report('count sheep')
      wolves = netlogo.report('count wolves')
      grass = netlogo.report('count grass')
      ticks = netlogo.report('ticks')
      comb_number = i
      # add to data frame
      parameter_space_df = parameter_space_df.append({'run': j, 'initial_number_sheep': initial_number_sheep, 'initial_number_wolves': initial_number_wolves, 'sheep_gain_from_food': sheep_gain_from_food, 'wolf_gain_from_food': wolf_gain_from_food, 'sheep_reproduce': sheep_reproduce, 'wolf_reproduce': wolf_reproduce, 'grass_regrowth': grass_regrowth, 'sheep': sheep, 'wolves': wolves, 'grass': grass, 'ticks': ticks, 'combination_number': comb_number}, ignore_index = True)
      # break out of this level if the maximum number of sheep is reached and there are no wolves or grass left
      if wolves == 0 and grass == 0 and sheep > 30000:
        break
      
  print(''.join(['finished run ', str(i), '\n']), flush = True)

# save data as csv file
parameter_space_df.to_csv("C:/Users/Femke Keij/OneDrive/Thesis/R projects/LU thesis local/data/raw/wolfsheep_output.csv")
```

Close the model:
```{python}
netlogo.kill_workspace()
```

Stop running the python environment:
```{python}
quit
```

# Split into training and test data
Clear working directory:
```{r}
rm(list = ls(all = TRUE)) 
```

Import csv output format for the model as tidy data. We ignore the first column because it contains the indices.
```{r}
ws_output <- read_csv("data/raw/wolfsheep_output.csv",
                        col_names = TRUE)[, - 1]
```

Manipulate data so that there is 1 outcome for each parameter combination:
The outcomes for each parameter set are summarised by their mean sheep, wolves, and grass at every tick of the simulation, as well as the min and max of those numbers.
```{r}
all_comb <- read_csv('data/processed/wolfsheep_parameterspace.csv')

# group the data by parameter combination and ticks
ws_summary <- ws_output %>%
  group_by(combination_number, ticks) %>%
  # take the mean, min, and max wolf, sheep, and grass numbers over the 15 runs for each density
  summarise(mean_sheep = mean(sheep),
            min_sheep = min(sheep),
            max_sheep = max(sheep),
            mean_wolves = mean(wolves),
            min_wolves = min(wolves),
            max_wolves = max(wolves),
            mean_grass = mean(grass),
            min_grass = min(grass),
            max_grass = max(grass),
            .groups = 'keep')

# add the initial parameters back in
ws_summary$initial_number_sheep <- numeric(nrow(ws_summary))
ws_summary$initial_number_wolves <- numeric(nrow(ws_summary))
ws_summary$sheep_gain_from_food <- numeric(nrow(ws_summary))
ws_summary$wolves_gain_from_food <- numeric(nrow(ws_summary))
ws_summary$sheep_reproduce <- numeric(nrow(ws_summary))
ws_summary$wolves_reproduce <- numeric(nrow(ws_summary))
ws_summary$grass_regrowth <- numeric(nrow(ws_summary))

# change the combination numbers from 0 - 39 to 1 - 40
ws_summary$combination_number <- ws_summary$combination_number + 1

for(i in 1:nrow(ws_summary)){
  comb_numb <- ws_summary$combination_number[i]
  ws_summary$initial_number_sheep[i] <- as.numeric(all_comb[comb_numb, 1])
  ws_summary$initial_number_wolves[i] <- as.numeric(all_comb[comb_numb, 2])
  ws_summary$sheep_gain_from_food[i] <- as.numeric(all_comb[comb_numb, 3])
  ws_summary$wolves_gain_from_food[i] <- as.numeric(all_comb[comb_numb, 4])
  ws_summary$sheep_reproduce[i] <- as.numeric(all_comb[comb_numb, 5])
  ws_summary$wolves_reproduce[i] <- as.numeric(all_comb[comb_numb, 6])
  ws_summary$grass_regrowth[i] <- as.numeric(all_comb[comb_numb, 7])
}
  
ws_summary %>%
  write_csv(file = 'data/processed/wolfsheep_summary.csv')
```

Set up the training and test data:
I'm using a 80 - 20 split.
```{r}
set.seed(42)

ind <- sample(1:nrow(ws_output),
              size = nrow(ws_output) * 0.2,
              replace = FALSE)

ws_output %>%
  slice(ind) %>%
  write_csv(file = 'data/processed/wolfsheep_test.csv')

ws_output %>%
  slice(-ind) %>%
  write_csv(file = 'data/processed/wolfsheep_train.csv')
```