---
title: "Try-out NetLogo R connectivity"
author: "Femke Keij S2647168"
date: '2022-07-12'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##NLXR package paper
Defining an`nl` object:
```{r}
library(nlrx)

nl <- nl(nlversion = "6.2.1",
         nlpath = "C://Program Files/Netlogo 6.2.1/",
         modelpath = "C://Program Files/Netlogo 6.2.1/app/models/Sample Models/Biology/Ants.nlogo",
         jvmmem = 1024)
```
An S4 class object that stores basic information on NetLogo. Contains information on
\item `nlversion`: the version
\item `nlpath`: the path to the NetLogo directory
\item `modelpath`: the path to the NetLogo model file
\item `jvmmem`: the amount of reserved memory for each Java virtual machine

Create an `experiment` and attach it to the `nl` object:
```{r}
nl@experiment <- experiment(expname = "nlrx_ants_school",
                            outpath = "/output",
                            repetition = 1,
                            tickmetrics = "false",
                            idsetup = "setup",
                            idgo = "go",
                            runtime = 0,
                            stopcond = "not any? patches with [food > 0]",
                            metrics = c("ticks"),
                            variables = list("population" = list(min = 10, max = 200, qfun = "qunif"),
                                             "diffusion-rate" = list(min = 1, max = 99, qfun = "qunif"),
                                             "evaporation-rate" = list(min = 1, max = 99, qfun = "qunif")
                                             )
)
``` 
Stores all information that would be entered into a typical BehaviorSpace experiment. The `experiment` object contains 
\item `expname`: the name of the experiment
\item `outpath`: the directory where output is written to
\item `repetition`: the number of repeated runs within BehaviorSpace
\item `tickmetrics`: a logical variable indicating whether output should be measured on each tick (`tickmetrics = "true"`) or on the final tick only (`tickemtrics = "false"`)
\item `idsetup` / `idgo`: names of setup and go procedures
\item `runtime`: the maximum number of ticks that should be simualted (0 = infinite)
\item `evalticks`: a vector of ticks for which output is reported
\item `stopcond`: a stop condition
\item `metrics` / `metrics.turtles` / `metrics.patches` / `metrics.links`: output metrics
\item `variables`: variable parameters and corresponding value ranges
\item `constants`: and constant parameters

A `simdesign` needs to be attached to the `nl` object in order to construct an input parameter matrix from the defined experiment
```{r}
nl@simdesign <- simdesign_sobol(nl = nl,
                                samples = 10,
                                sobolorder = 2,
                                sobolnboot = 100,
                                sobolconf = 0.95,
                                nseeds = 6,
                                precision = 3)
```
The NLRX package provides many helper functions to create pre-designed `simdesigns` for simple parameter screenings (simple, distinct, full-factorial), sensitivity analyses, or parameter optimizations (simulated annealing, genetic algorithm). All these helper functions take information from the experiment (`variables`, `constants`) to create a `simdesign` object that contains
\item `simmethod`: the name of the simulation design method
\item `simseeds`: a vector of random seeds
\item `siminput`: the generated parameter matrix
\item `simoutput`: an empty tibble to slot to attach output results after simulations have been finished
\item `simobject`: the sensitivity and optimization simdesign helper functions also create this object, which contains further information needed for post-processing, such as calculation of sensitivity indices

After initialization of the `nl` object, simulations can be run by calling one of the `run_nl_*()` functions
```{r}
results <- run_nl_all(nl)
# there is only 1 result, because we asked the experiment to only report the final simulation step
```
There are 3 options:
\item `run_nl_one()`: for `simdesigns` that generate a parameter matrix, this function can be used to execute one specific row of the `simdesign` parameter matrix `siminput` by specifying the row ID and model seed
\item `run_nl_all()`: executes all simulations from the parameter matrix `siminput` across all seeds
\item `run_nl_dyn()`: for optimization `simdesigns` that do not come with a pre-generated parameter input matrix. This executes the model simulations with dynamically generated parameter settings (i.e., when the parameter settings of the next run depend on those of previous runs)

After simulations are finished, the resulting output table can be attached to the `nl` object by using the function `setsim`
```{r}
setsim(nl, "simoutput") <- results
```
This stores a complete experiment setup, including output results, in one R object.

The output can also be written to a .csv file using `write_simoutput` and further anlayzed using `analyze_nl()`.
```{r}
sensitivityIndices <- analyze_nl(nl)

write_simoutput(nl)
```

Other options
\item `unnest_simoutput()`: to turn agent-specific metrics collected during the simulation into a wide-format table
\item `nl_to_points()` / `nl_to_raster()` / `nl_to_graph()`: to coerce the `nl` object into a spatial data type
\item `export_nl()` / `import_nl()`: to store R and corresponding NetLogo model scripts in a zip file, which uses relative paths and thus enables easy collaboration

### Optimization with a Genetic Algorithm

## Let's try RNetLogo with NetLogo 5.0.5
Downloaded NetLogo 5.0.5 (version I had was 6.2.1) to see if I can get RNetLogo to work with that version. Following tutorials from LU40 Thiele et al. 2014.

Start by creating a NetLogo instance by using `NLStart`. Then load a model with `NLLoadModel` and then use commands and reports to interact with it.

```{r}
# load the package
library(RNetLogo)

# initialize NetLogo
nl.path <- "C:/Program Files/Netlogo 6.0.4"
nl.jarname <- "netlogo-6.0.4.jar"
#nl.jarname <- "Netlogo.jar"
NLStart(nl.path = nl.path, gui = FALSE, nl.jarname = nl.jarname)

NLLoadModel("/models/Sample Models/Earth Science/Fire.nlogo")

# load the NetLogo model
model.path <- "C:/models/woodhoopoe.nlogo"
NLLoadModel(model.path)


nl.path <- "C:/Program Files/NetLogo 6.0.4/app"
nl.jarname <- "netlogo-6.0.4.jar"
NLStart(nl.path = nl.path, gui = FALSE, nl.jarname = nl.jarname)
NLLoadModel("C:/0Werk/Onderwijs/2019/06ESSA-Aberdeen/Rdata_Files/Fire.nlogo")
```
