---
title: "Fire_07_Calibration_RandomForest"
author: "Femke Keij S2647168"
date: "2023-02-14"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preliminaries
Clear working directory & set random seed:
```{r}
rm(list = ls(all = TRUE))

set.seed(42)
```

Packages used:
```{r}
# for importing / working with tidy data
library(tidyverse)

# for ggplot
library(directlabels)
library(patchwork)
library(ggbeeswarm)

# for fitting the random forest
library(randomForest)
```

Source files for error computation and plotting functions:
```{r}
source('scripts/Fire model/Fire_ErrorFunctions.R')
```

Read in the data:
```{r}
# complete output data
fire_output <- read_csv('data/raw/fire_output.csv')

# training data
fire_train <- read_csv(
  'data/processed/fire_train.csv')
fire_train$directions <- as.factor(fire_train$directions)

# test data
fire_test <- read_csv(
  'data/processed/fire_test.csv')
fire_test$directions <- as.factor(fire_test$directions)
```

Random samples:
5, 10, 20, 30, 40, 50, 60, 70, and 80 $%$ of the total parameter space.
```{r}
N <- nrow(fire_output)
# range of (training) sample sizes to work with
n <- c(0.05 * N, 0.1 * N, 0.2 * N, 0.3 * N, 0.4 * N, 0.5 * N,
       0.6 * N, 0.7 * N, nrow(fire_train))
```

# Fitting
Here I'm fitting a random forest to predict density and direction.
```{r}
RandomForestFitting <- function(ticks = FALSE){
  # table to store all output
  length_tib <- sum(nrow(fire_test) * length(n))
  forest_results <- tibble(direction_true = numeric(length_tib),
                              direction_pred = numeric(length_tib),
                              density_true = numeric(length_tib),
                              density_pred = numeric(length_tib),
                              burn_true = numeric(length_tib),
                              burn_pred = numeric(length_tib),
                              n = numeric(length_tib),
                              ticks_included = numeric(length_tib))
  # vector to store mean density in each training sample
  # (to calculate point prediction performance later on)
  density_means <- numeric(length(n))
  # list to store fitted Elastic Nets
  forest_fits <- list()

  count <- 1
  
  # for each sample size
  for(i in 1:length(n)){
    # sample training set
    ind_sample <- sample(1:nrow(fire_train), size = n[i])
    fire_train_sub <- fire_train[ind_sample, ]
    # calculate and save mean density in training set
    density_means[i] <- mean(fire_train_sub$density)
    
    if(ticks == TRUE){
      # random forest to predict density when ticks are included
      fit_forest_density <- randomForest(density ~ burn_percentage +
                                            directions + ticks,
                                          data = fire_train_sub)
      # random forest to predict directions when ticks are included
      fit_forest_directions <- randomForest(directions ~ burn_percentage +
                                               density + ticks,
                                             data = fire_train_sub)
      
    } else{
      # random forest to predict density when ticks are excluded
      fit_forest_density <- randomForest(density ~ burn_percentage +
                                           directions,
                                         data = fire_train_sub)
      # random forest to predict directions when ticks are excluded
      fit_forest_directions <- randomForest(directions ~ burn_percentage +
                                              density,
                                            data = fire_train_sub)
    }
    # predict density in test data using fitted random forest
    predict_forest_density <- predict(fit_forest_density,
                                       newdata = fire_test)
    # predict direction in test data using fitted random forest
    predict_forest_directions <- predict(fit_forest_directions,
                                          newdata = fire_test)
    
    # for each instance of the test data
    for(j in 1:nrow(fire_test)){
      # find an instance in the training data that matches the directions
      # and density
      ind <- which(fire_train$density == round(predict_forest_density[j]) &
                     fire_train$directions ==
                     predict_forest_directions[j])
      # if there are multiple, pick one
      if(length(ind) > 1){
        ind <- sample(ind, size = 1)
        # if there are none, increase density by 1 and see if there
        # is a matching instance in the training data
        # if not, continue until a matching instance is found
      } else if (length(ind) == 0){
        add <- 1
        while(length(ind) != 1){
          if(round(predict_forest_density[j]) > 50){
            add <- add * -1
          }
          ind <- which(fire_train$density ==
                         (round(predict_forest_density[j]) + add) &
                     fire_train$directions ==
                     predict_forest_directions[j])
          if(length(ind) > 1){
            ind <- sample(ind, size = 1)
          }
          add <- abs(add) + 1
        }
      }
      # store all results in table
      forest_results$direction_true[count] <- fire_test$directions[j]
      forest_results$direction_pred[count] <-
        predict_forest_directions[j]
      forest_results$density_true[count] <- fire_test$density[j]
      forest_results$density_pred[count] <- predict_forest_density[j]
      forest_results$burn_true[count] <- fire_test$burn_percentage[j]
      forest_results$burn_pred[count] <- fire_train$burn_percentage[ind]
      forest_results$n[count] <- n[i]
      forest_results$ticks_included[count] <- if(ticks){'yes'}else{'no'}
      
      count <- count + 1
    }
    # store regression fit
    forest_fits <- append(forest_fits, fit_forest_density)
    forest_fits <- append(forest_fits, fit_forest_directions)
    
    print(paste('finished fitting sample size ', n[i]))
  }
  
  # correct format of direction_true variable
  forest_results %>%
    mutate(direction_true = replace(direction_true, direction_true == 1, 4),
         direction_true = replace(direction_true, direction_true == 2, 8),
         direction_pred = replace(direction_pred, direction_pred == 1, 4),
         direction_pred = replace(direction_pred, direction_pred == 2, 8)) ->
    forest_results
  
  # return list with results, density means, and fitted random forests
  return(list(forest_results, density_means, forest_fits))
}
```

```{r}
# fit regressions with ticks
forest_results_ticks <- RandomForestFitting(ticks = TRUE)
density_means_ticks <- forest_results_ticks[[2]]
forest_fits_ticks <- forest_results_ticks[[3]]
forest_results_ticks <- forest_results_ticks[[1]]

# fit regressions without ticks
forest_results_noticks <- RandomForestFitting(ticks = FALSE)
density_means_noticks <- forest_results_noticks[[2]]
forest_fits_noticks <- forest_results_noticks[[3]]
forest_results_noticks <- forest_results_noticks[[1]]
```

```{r}
# run error calculations
forest_errors_ticks <- ComputeErrors(forest_results_ticks,
                                         ticks = TRUE,
                                         density_means_ticks,
                                         pred_int_included = FALSE)
forest_errors_noticks <- ComputeErrors(forest_results_noticks,
                                           ticks = FALSE,
                                           density_means_noticks,
                                           pred_int_included = FALSE)
```

Merge results
```{r}
forest_results <- rbind(forest_results_noticks,
                            forest_results_ticks)
write_csv(forest_results, 'data/processed/fire_forest_results.csv')

forest_errors <- rbind(forest_errors_noticks, forest_errors_ticks)
write_csv(forest_errors, 'data/processed/fire_forest_errors.csv')
``` 

# Plots
```{r}
# forest_results <- read_csv('data/processed/fire_forest_results.csv')
# forest_errors <- read_csv('data/processed/fire_forest_errors.csv')
```

## Errors
Fix results and errors data frames:
```{r}
forest_results <- forest_results %>%
  mutate(n = factor(n),
         direction_true = factor(direction_true),
         direction_pred = factor(direction_pred),
         direction_correct = (direction_pred == direction_true))

forest_errors <- forest_errors %>%
  mutate(n = factor(n))
```

### Predicted vs. true
Predicted vs. true for density, directions, and burn percentage
```{r}
p1 <- PlotPredTrueDensity(forest_results, n = 1584)
p2 <- PlotPredTrueDirection(forest_results, n = 1584)
p3 <- PlotPredTrueBurn(forest_results, n = 1584)

p1 / p2 / p3
```
Predicted vs. true density (largest sample size only):
```{r}
PlotPredTrueDensity(forest_results, n = 1584)

ggsave('figures/fire_randomforest_predtrue_density.pdf')
```
Confusion matrix for predicted vs. true directions:
```{r}
PlotPredTrueDirection(forest_results, n = 1584)

ggsave('figures/fire_randomforest_predtrue_direction.pdf')
```
Predicted vs. true burn percentage (largest sample size only):
```{r}
PlotPredTrueBurn(forest_results, n = 1584)

ggsave('figures/fire_randomforest_predtrue_burn.pdf')
```
### Parameters
% correctly predicted parameters (density & directions):
```{r}
PlotPercCorrectParams(forest_errors, n_plot = c(99, 792, 1584))

ggsave('figures/fire_randomforest_perc_correct.pdf')
```
RMSE, NRMSE, point prediction performance, and coverage for density parameter:
```{r}
PlotErrorsDensity(forest_errors, n_plot = c(99, 792, 1584))

ggsave('figures/fire_randomforest_density_metrics.pdf')
```
F1, kappa score, and MCC for directions:
```{r}
PlotErrorsDirections(forest_errors, n_plot = c(99, 792, 1584))

ggsave('figures/fire_randomforest_directions_metrics.pdf')
```
### Outcome variable
RMSE for burn percentage:
```{r}
PlotRMSEOut(forest_errors, n_plot = c(99, 792, 1584))

ggsave('figures/fire_randomforest_RMSE_burn.pdf')
```

## Model fit
Variable importance plots for density & directions
```{r}
fit_forest_density_noticks <- randomForest(density ~ burn_percentage +
                                            directions,
                                          data = fire_train,
                                   importance = TRUE)

imp_noticks <- as.data.frame(varImpPlot(fit_forest_density_noticks))
imp_noticks$varNames <- rownames(imp_noticks) # row names to column
colnames(imp_noticks) <- c('percIncMSE', 'IncNodePurity', 'varNames')

p1 <- imp_noticks %>%
  ggplot(aes(x = varNames,
             weight = percIncMSE, y = percIncMSE,
             colour = varNames)) +
  geom_point(size = 3) +
  geom_segment(aes(x = varNames, xend = varNames, y = 0, yend = percIncMSE),
               linewidth = 1.5) +
  labs(y = '% increase MSE when removing variable',
       x = NULL) +
  ylim(0, 50) +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
        panel.border = element_rect(colour = 'lightgrey', fill = NA),
        legend.position = 'none') +
  ggtitle('Density, no ticks')

fit_forest_density_ticks <- randomForest(density ~ burn_percentage +
                                            directions + ticks,
                                          data = fire_train,
                                   importance = TRUE)

imp_ticks <- as.data.frame(varImpPlot(fit_forest_density_ticks))
imp_ticks$varNames <- rownames(imp_ticks) # row names to column
colnames(imp_ticks) <- c('percIncMSE', 'IncNodePurity', 'varNames')

p2 <- imp_ticks %>%
  ggplot(aes(x = varNames,
             weight = percIncMSE, y = percIncMSE,
             colour = varNames)) +
  geom_point(size = 3) +
  geom_segment(aes(x = varNames, xend = varNames, y = 0, yend = percIncMSE),
               linewidth = 1.5) +
  labs(y = '% increase MSE when removing variable',
       x = NULL) +
  ylim(0, 50) +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
        panel.border = element_rect(colour = 'lightgrey', fill = NA),
        legend.position = 'none') +
  ggtitle('Density, ticks')

fit_forest_directions_noticks <- randomForest(directions ~ burn_percentage +
                                            density,
                                          data = fire_train,
                                   importance = TRUE)

imp_directions_noticks <-
  as.data.frame(varImpPlot(fit_forest_directions_noticks))
imp_directions_noticks$varNames <- 
  rownames(imp_directions_noticks) # row names to column
colnames(imp_directions_noticks) <- c('percIncMSE', 'IncNodePurity',
                                      'varNames')

p3 <- imp_directions_noticks %>%
  ggplot(aes(x = varNames,
             weight = percIncMSE, y = percIncMSE,
             colour = varNames)) +
  geom_point(size = 3) +
  geom_segment(aes(x = varNames, xend = varNames, y = 0, yend = percIncMSE),
               linewidth = 1.5) +
  labs(y = '% increase MSE when removing variable',
       x = NULL) +
  ylim(0, 375) +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
        panel.border = element_rect(colour = 'lightgrey', fill = NA),
        legend.position = 'none') +
  ggtitle('Directions, no ticks')

fit_forest_directions_ticks <- randomForest(directions ~ burn_percentage +
                                            density + ticks,
                                          data = fire_train,
                                   importance = TRUE)

imp_directions_ticks <- as.data.frame(varImpPlot(fit_forest_directions_ticks))
imp_directions_ticks$varNames <- 
  rownames(imp_directions_ticks) # row names to column
colnames(imp_directions_ticks) <- c('percIncMSE', 'IncNodePurity', 'varNames')

p4 <- imp_directions_ticks %>%
  ggplot(aes(x = varNames,
             weight = percIncMSE, y = percIncMSE,
             colour = varNames)) +
  geom_point(size = 3) +
  geom_segment(aes(x = varNames, xend = varNames, y = 0, yend = percIncMSE),
               linewidth = 1.5) +
  labs(y = '% increase MSE when removing variable',
       x = NULL) +
  ylim(0, 375) +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
        panel.border = element_rect(colour = 'lightgrey', fill = NA),
        legend.position = 'none') +
  ggtitle('Directions, ticks')

(p1 + p2) / (p3 + p4)
  plot_layout(guides = 'collect') +
  plot_annotation(tag_levels = 'A') &
  theme(plot.tag = element_text(size = 8))
  
ggsave('figures/fire_forest_varimportance.pdf')
```

I could still add partial dependence plots if I want to.